---
title: "Data set creation"
author: "Iona Lyell"
date: "19/08/2021"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### load packages
```{r}
library(readr)
library(readxl)
library(data.table)
library(ggplot2)
library(dplyr)
library(skimr)
```



## GENDER SOCIAL NORMS INDEX

from: http://hdr.undp.org/en/gsni  Gender Social Norms Index data tables (Excel)

```{r}
gsni_tables <- read_excel("data/gsni_tables.xlsx")
gsni_tables <- as.data.table(gsni_tables)

#rename columns
setnames(gsni_tables, "Table A1: Gender Social Norms Index (GSNI), last available period", "location_name")
setnames(gsni_tables, "...2", "GSNI_PERIOD")
setnames(gsni_tables, "...3", "onebias") #share of people with at least 1 bias (%)
setnames(gsni_tables, "...4", "twobias") #share of people with at least 2 biases (%)
setnames(gsni_tables, "...5", "nobias") #share of people with no (%)
setnames(gsni_tables, "...6", "political") #share of people with political bias (%)
setnames(gsni_tables, "...7", "economic") #share of people with economic bias (%)
setnames(gsni_tables, "...8", "educational") #share of people with educational bias (%)
setnames(gsni_tables, "...9", "physical") #share of people with physical bias (%)


gsni_tables <- gsni_tables[8:82]
 #n = 75


```

## CVD MORTALITY RATES PER COUNTRY FOR MEN and WOMEN AND OVERALL 

from: http://ghdx.healthdata.org/gbd-results-tool

Measure: Deaths
Age: Age standardized   
Year: 2005-2019 (data collections for GINI 2005-2009 or 2010-2014) 
Need to use all the same year probably?2014
Cause: B.2 cardiovascular disease
Context: Cause
Location: Select only countries and territories
Sex: Male + Female + Both
Metric: Rate

Deaths per 100, 000 population (confimred by GBD CODEBOOK IHME_GBD_2019_MEASURE_METRIC_DEFINITIONS_Y2020M10D15) from http://ghdx.healthdata.org/gbd-results-tool

```{r}
gbd_cvd_mort <- read_csv("data/IHME-GBD_2019_DATA-703c6c0a-1/IHME-GBD_2019_DATA-703c6c0a-1.csv")
gbd_cvd_mort <- as.data.table(gbd_cvd_mort)


```
Create columns of each sex catergory and year...
```{r}
gbd_cvd_mort <- gbd_cvd_mort[, .(location_name, sex_name, year, val)] #just columns we want

gbd_cvd_mort_split <- split(gbd_cvd_mort, gbd_cvd_mort$year)

cvd_2005 <- gbd_cvd_mort_split$`2005`
cvd_2006 <- gbd_cvd_mort_split$`2006`
cvd_2007 <- gbd_cvd_mort_split$`2007`
cvd_2008 <- gbd_cvd_mort_split$`2008`
cvd_2009 <- gbd_cvd_mort_split$`2009`
cvd_2010 <- gbd_cvd_mort_split$`2010`
cvd_2011 <- gbd_cvd_mort_split$`2011`
cvd_2012 <- gbd_cvd_mort_split$`2012`
cvd_2013 <- gbd_cvd_mort_split$`2013`
cvd_2014 <- gbd_cvd_mort_split$`2014`
cvd_2015 <- gbd_cvd_mort_split$`2015`
cvd_2016 <- gbd_cvd_mort_split$`2016`
cvd_2017 <- gbd_cvd_mort_split$`2017`
cvd_2018 <- gbd_cvd_mort_split$`2018`
cvd_2019 <- gbd_cvd_mort_split$`2019`
setnames(cvd_2005, "val", "cvd2005")
setnames(cvd_2006, "val", "cvd2006")
setnames(cvd_2007, "val", "cvd2007")
setnames(cvd_2008, "val", "cvd2008")
setnames(cvd_2009, "val", "cvd2009")
setnames(cvd_2010, "val", "cvd2010")
setnames(cvd_2011, "val", "cvd2011")
setnames(cvd_2012, "val", "cvd2012")
setnames(cvd_2013, "val", "cvd2013")
setnames(cvd_2014, "val", "cvd2014")
setnames(cvd_2015, "val", "cvd2015")
setnames(cvd_2016, "val", "cvd2016")
setnames(cvd_2017, "val", "cvd2017")
setnames(cvd_2018, "val", "cvd2018")
setnames(cvd_2019, "val", "cvd2019")
cvd_2005[, year:= NULL]
cvd_2006[, year:= NULL]
cvd_2007[, year:= NULL]
cvd_2008[, year:= NULL]
cvd_2009[, year:= NULL]
cvd_2010[, year:= NULL]
cvd_2011[, year:= NULL]
cvd_2012[, year:= NULL]
cvd_2013[, year:= NULL]
cvd_2014[, year:= NULL]
cvd_2015[, year:= NULL]
cvd_2016[, year:= NULL]
cvd_2017[, year:= NULL]
cvd_2018[, year:= NULL]
cvd_2019[, year:= NULL]


MyMerge <- function(x, y){
  df <- merge(x, y, by= c("location_name", "sex_name"), all.x= TRUE, all.y= TRUE)
  return(df)
}
cvdallyears <- Reduce(MyMerge, list(cvd_2005, cvd_2006, cvd_2007, cvd_2008, cvd_2009, cvd_2010, cvd_2011, cvd_2012, cvd_2013, cvd_2014, cvd_2015, cvd_2016, cvd_2017, cvd_2018,cvd_2019))


cvdallyears_split <- split(cvdallyears, cvdallyears$sex_name)
cvdallyearsboth <- cvdallyears_split$`Both`
cvdallyearsfemale <- cvdallyears_split$`Female`
cvdallyearsmale <- cvdallyears_split$`Male`

names(cvdallyearsboth)[3:17] <- c("cvd05both", "cvd06both", "cvd07both", "cvd08both", "cvd09both", "cvd10both", "cvd11both", "cvd12both", "cvd13both", "cvd14both", "cvd15both", "cvd16both", "cvd17both", "cvd18both", "cvd19both")
cvdallyearsboth[, sex_name := NULL]

names(cvdallyearsfemale)[3:17] <- c("cvd05fem", "cvd06fem", "cvd07fem", "cvd08fem", "cvd09fem", "cvd10fem", "cvd11fem", "cvd12fem", "cvd13fem", "cvd14fem", "cvd15fem", "cvd16fem", "cvd17fem", "cvd18fem", "cvd19fem")
cvdallyearsfemale[, sex_name := NULL]

names(cvdallyearsmale)[3:17] <- c("cvd05male", "cvd06male", "cvd07male", "cvd08male", "cvd09male", "cvd10male", "cvd11male", "cvd12male", "cvd13male", "cvd14male", "cvd15male", "cvd16male", "cvd17male", "cvd18male", "cvd19male")
cvdallyearsmale[, sex_name := NULL]

MyMerge <- function(x, y){
  df <- merge(x, y, by= c("location_name"), all.x= TRUE, all.y= TRUE)
  return(df)
}
cvdfinal<- Reduce(MyMerge, list(cvdallyearsfemale, cvdallyearsmale, cvdallyearsboth))

```



```{r}

#rename countries

cvdfinal$location_name[cvdfinal$location_name == "Iran (Islamic Republic of)"] <- "Iran, Islamic Republic of"

cvdfinal$location_name[cvdfinal$location_name == "Republic of Korea"] <- "Korea (Republic of)"
#Democratic People's Republic of Korea = north korea, Republic of Korea = south korea Korea (republic of) is in gsni tables

cvdfinal$location_name[cvdfinal$location_name == "Republic of Moldova"] <- "Moldova, Republic of"

cvdfinal$location_name[cvdfinal$location_name == "Palestine"] <- "Palestine, State of"

cvdfinal$location_name[cvdfinal$location_name == "United States of America"] <- "United States"

gsni_cvd <- merge(gsni_tables, cvdfinal, by = "location_name", all = TRUE) #combine gsni and cvd mort for both sexes

gsni_cvd <- na.omit(gsni_cvd, cols = "GSNI_PERIOD") #omit countries without GSNI and cvd mort


#n=75

```




## LIFE EXPECTANCY FOR WOMEN

Per country for women at birth and age 60 (?use age 60 to remove confounding of MMR)
2000, 2010, 2015, 2019

from: https://apps.who.int/gho/data/view.main.WOMENLEXv

```{r}
fle <- read_csv("data/female_life_expectancy.csv")
fle <- as.data.table(fle)
fle <- fle[ -1,]

names(fle) <- c("location_name", "year", "LEbirth", "LE60")



female_life_expectancy_split <- split(fle, fle$year)


FLE2000 <- female_life_expectancy_split$`2000`
setnames(FLE2000, old = c("LEbirth", "LE60"), new = c("LEbirth2000", "LE602000"))
FLE2000[, year:= NULL]

FLE2010 <- female_life_expectancy_split$`2010`
setnames(FLE2010, old = c("LEbirth", "LE60"), new = c("LEbirth2010", "LE602010"))
FLE2010[, year:= NULL]

FLE2015 <- female_life_expectancy_split$`2015`
setnames(FLE2015, old = c("LEbirth", "LE60"), new = c("LEbirth2015", "LE602015"))
FLE2015[, year:= NULL]

FLE2019 <- female_life_expectancy_split$`2019`
setnames(FLE2019, old = c("LEbirth", "LE60"), new = c("LEbirth2019", "LE602019"))
FLE2019[, year:= NULL]


MyMerge <- function(x, y){
  df <- merge(x, y, by= c("location_name"), all.x= TRUE, all.y= TRUE)
  return(df)
}
female_life_expectancy<- Reduce(MyMerge, list(FLE2000, FLE2010, FLE2015, FLE2019))


```

```{r}

#rename countries for merge
female_life_expectancy$location_name[female_life_expectancy$location_name == "Iran (Islamic Republic of)"] <- "Iran, Islamic Republic of"

female_life_expectancy$location_name[female_life_expectancy$location_name == "Republic of Korea"] <- "Korea (Republic of)"
#Democratic People's Republic of Korea = north korea, Republic of Korea = south korea Korea (republic of) is in gsni tables 

female_life_expectancy$location_name[female_life_expectancy$location_name == "Republic of Moldova"] <- "Moldova, Republic of"

female_life_expectancy$location_name[female_life_expectancy$location_name == "United States of America"] <- "United States"


gsni_cvd_lifexp <- merge(gsni_cvd, female_life_expectancy, by = "location_name", all = TRUE) 
gsni_cvd_lifexp <- na.omit(gsni_cvd_lifexp, cols = "GSNI_PERIOD") #omit countries without GSNI

```

## LIFE EXPECTANCY FOR MEN

Per country for men at birth
2000, 2010, 2015, 2019

from: https://apps.who.int/gho/data/view.main.SDG2016LEXv

```{r}
mle <- read_csv("data/WHO_LE.csv")
mle <- as.data.table(mle)
mle <- mle[,c(1,2,4)]
mle <- mle[-1,]

names(mle) <- c("location_name", "year", "MLEbirth")



male_life_expectancy_split <- split(mle, mle$year)


mle2000 <- male_life_expectancy_split$`2000`
setnames(mle2000, old = c("MLEbirth"), new = c("MLEbirth2000"))
mle2000[, year:= NULL]

mle2010 <- male_life_expectancy_split$`2010`
setnames(mle2010, old = c("MLEbirth"), new = c("MLEbirth2010"))
mle2010[, year:= NULL]

mle2015 <- male_life_expectancy_split$`2015`
setnames(mle2015, old = c("MLEbirth"), new = c("MLEbirth2015"))
mle2015[, year:= NULL]

mle2019 <- male_life_expectancy_split$`2019`
setnames(mle2019, old = c("MLEbirth"), new = c("MLEbirth2019"))
mle2019[, year:= NULL]


MyMerge <- function(x, y){
  df <- merge(x, y, by= c("location_name"), all.x= TRUE, all.y= TRUE)
  return(df)
}
male_life_expectancy<- Reduce(MyMerge, list(mle2000, mle2010, mle2015, mle2019))


```

```{r}

#rename countries for merge
male_life_expectancy$location_name[male_life_expectancy$location_name == "Iran (Islamic Republic of)"] <- "Iran, Islamic Republic of"

male_life_expectancy$location_name[male_life_expectancy$location_name == "Republic of Korea"] <- "Korea (Republic of)"
#Democratic People's Republic of Korea = north korea, Republic of Korea = south korea Korea (republic of) is in gsni tables 

male_life_expectancy$location_name[male_life_expectancy$location_name == "Republic of Moldova"] <- "Moldova, Republic of"

male_life_expectancy$location_name[male_life_expectancy$location_name == "United States of America"] <- "United States"


gsni_cvd_lifexpmf <- merge(gsni_cvd_lifexp, male_life_expectancy, by = "location_name", all = TRUE) 
gsni_cvd_lifexpmf <- na.omit(gsni_cvd_lifexpmf, cols = "GSNI_PERIOD") #omit countries without GSNI

```

## PHYSICIANS PER 1000

from: https://data.worldbank.org/indicator/SH.MED.PHYS.ZS

Didn't work when downloaded as csv but worked when downloaded as xlsx
row 4 is year, data from 1960-2020
last updated 17/02/2021

want 2005-2017

```{r}
phys_per_1000 <- read_excel("data/physiciansper1000/API_SH.MED.PHYS.ZS_DS2_en_csv_v2_2063402.xlsx")
phys_per_1000 <- as.data.table(phys_per_1000)

#2005-2017 (very little data for 18/19) columns...50-...62

phys_05_17 <- phys_per_1000 %>% select(1,50:62)
phys_05_17 <- phys_05_17[-c(1:4),]

names(phys_05_17) <- c("location_name", "phy05", "phy06", "phy07", "phy08", "phy09", "phy10", "phy11", "phy12", "phy13", "phy14", "phy15", "phy16", "phy17")

```

```{r}

#rename countries for merge
phys_05_17$location_name[phys_05_17$location_name == "Iran, Islamic Rep."] <- "Iran, Islamic Republic of"

phys_05_17$location_name[phys_05_17$location_name == "Korea, Rep."] <- "Korea (Republic of)"
#Democratic People's Republic of Korea = north korea, Republic of Korea = south korea Korea (republic of) is in gsni tables

phys_05_17$location_name[phys_05_17$location_name == "Moldova"] <- "Moldova, Republic of"

gsni_cvd_lifexp_phys <- merge(gsni_cvd_lifexpmf, phys_05_17, by = "location_name", all = TRUE) 
gsni_cvd_lifexp_phys <- na.omit(gsni_cvd_lifexp_phys, cols = "GSNI_PERIOD") #omit countries without GSNI

```


## HEALTH EXPENDITURE (% of GDP)

from: https://data.worldbank.org/indicator/SH.XPD.CHEX.GD.ZS

```{r}
HE <- read_excel("data/API_SH.XPD.CHEX.GD.ZS_DS2_en_excel_v2_2445365.xls")


HE1 <- HE[, c(1,50:62)]


names(HE1) <- c("location_name", "HE05", "HE06", "HE07", "HE08", "HE09", "HE10", "HE11", "HE12", "HE13", "HE14", "HE15", "HE16", "HE17")


#rename countries for merge
HE1$location_name[HE1$location_name == "Iran, Islamic Rep."] <- "Iran, Islamic Republic of"

HE1$location_name[HE1$location_name == "Korea, Rep."] <- "Korea (Republic of)"
#Democratic People's Republic of Korea = north korea, Republic of Korea = south korea Korea (republic of) is in gsni tables

HE1$location_name[HE1$location_name == "Moldova"] <- "Moldova, Republic of"

gsni_cvd_lifexp_phys <- merge(gsni_cvd_lifexp_phys, HE1, by = "location_name", all = TRUE) 
gsni_cvd_lifexp_phys <- na.omit(gsni_cvd_lifexp_phys, cols = "GSNI_PERIOD") #omit countries without GSNI

```



## MEAN YEARS OF SCHOOLING

from: https://ourworldindata.org/global-education
up to 2017
want 2005-2017

```{r}
mean_years_school <- read_csv("data/mean-years-of-schooling-1.csv")
mean_years_school <- as.data.table(mean_years_school)

school_split <- split(mean_years_school, mean_years_school$Year)


scl2005 <- school_split$`2005`
names(scl2005) <- c("location_name","code", "year", "scl05")
scl2005 <- scl2005 %>% select(1,4)

scl2006 <- school_split$`2006`
names(scl2006)<- c("location_name","code", "year","scl06")
scl2006 <- scl2006 %>% select(1,4)

scl2007 <- school_split$`2007`
names(scl2007)<- c("location_name","code", "year","scl07")
scl2007 <- scl2007 %>% select(1,4)

scl2008 <- school_split$`2008`
names(scl2008)<- c("location_name","code", "year","scl08")
scl2008 <- scl2008 %>% select(1,4)

scl2009 <- school_split$`2009`
names(scl2009)<- c("location_name","code", "year","scl09")
scl2009 <- scl2009 %>% select(1,4)

scl2010 <- school_split$`2010`
names(scl2010)<- c("location_name","code", "year","scl10")
scl2010 <- scl2010 %>% select(1,4)

scl2011 <- school_split$`2011`
names(scl2011)<- c("location_name","code", "year","scl11")
scl2011 <- scl2011 %>% select(1,4)

scl2012 <- school_split$`2012`
names(scl2012)<- c("location_name","code", "year","scl12")
scl2012 <- scl2012 %>% select(1,4)

scl2013 <- school_split$`2013`
names(scl2013)<- c("location_name","code", "year","scl13")
scl2013 <- scl2013 %>% select(1,4)

scl2014 <- school_split$`2014`
names(scl2014) <- c("location_name", "code", "year", "scl14")
scl2014 <- scl2014 %>% select(1,4)

scl2015 <- school_split$`2015`
names(scl2015)<- c("location_name","code", "year","scl15")
scl2015 <- scl2015 %>% select(1,4)

scl2016 <- school_split$`2016`
names(scl2016)<- c("location_name","code", "year","scl16")
scl2016 <- scl2016 %>% select(1,4)

scl2017 <- school_split$`2017`
names(scl2017)<- c("location_name","code", "year","scl17")
scl2017 <- scl2017 %>% select(1,4)

MyMerge <- function(x, y){
  df <- merge(x, y, by= c("location_name"), all.x= TRUE, all.y= TRUE)
  return(df)
}
scl0517 <- Reduce(MyMerge, list(scl2005, scl2006, scl2007, scl2008, scl2009, scl2010, scl2011, scl2012, scl2013, scl2014, scl2015, scl2016, scl2017))


```

```{r}

#rename columns 
mean_years_school$location_name[mean_years_school$location_name == "Iran"] <- "Iran, Islamic Republic of"
mean_years_school$location_name[mean_years_school$location_name == "South Korea"] <- "Korea (Republic of)"

mean_years_school$location_name[mean_years_school$location_name == "Moldova"] <- "Moldova, Republic of"

mean_years_school$location_name[mean_years_school$location_name == "Palestine"] <- "Palestine, State of"


gsni_cvd_lifexp_phys_sch <- merge(gsni_cvd_lifexp_phys, scl0517, by = "location_name", all = TRUE) 
gsni_cvd_lifexp_phys_sch <- na.omit(gsni_cvd_lifexp_phys_sch, cols = "GSNI_PERIOD") #omit countries without GSNI

```

## GDP

GDP per capita (current US dollar)

from: https://data.worldbank.org/indicator/NY.GDP.PCAP.CD

```{r}
library(readxl)
GDPpercapita <- read_excel("data/GDPpercapita.xls")
GDPpercapita <- as.data.table(GDPpercapita)


#2005-2017 

GDP0517 <- GDPpercapita %>% select(1,50:62)
GDP0517 <- GDP0517[-c(1:3),]

names(GDP0517) <- c("location_name", "GDP05", "GDP06", "GDP07", "GDP08", "GDP09", "GDP10", "GDP11", "GDP12", "GDP13", "GDP14", "GDP15", "GDP16", "GDP17")


```

```{r}

#rename columns 
GDPpercapita$location_name[GDPpercapita$location_name == "Iran, Islamic Rep."] <- "Iran, Islamic Republic of"

GDPpercapita$location_name[GDPpercapita$location_name == "Korea, Rep."] <- "Korea (Republic of)"

GDPpercapita$location_name[GDPpercapita$location_name == "Moldova"] <- "Moldova, Republic of"


gsni_cvd_lifexp_phys_sch_gdp <- merge(gsni_cvd_lifexp_phys_sch, GDP0517, by = "location_name", all = TRUE) 
gsni_cvd_lifexp_phys_sch_gdp <- na.omit(gsni_cvd_lifexp_phys_sch_gdp, cols = "GSNI_PERIOD") #omit countries without GSNI



```






## MATERNAL MORTALITY RATIO

from: https://data.worldbank.org/indicator/SH.STA.MMRT

```{r}
MMR <- read_excel("data/API_SH.STA.MMRT_DS2_en_excel_v2_2056748.xls")
MMR <- as.data.table(MMR)

#2005-2017 

MMR0517 <- MMR %>% select(1,50:62)
MMR0517 <- MMR0517[-c(1:3),]

names(MMR0517) <- c("location_name", "MMR05", "MMR06", "MMR07", "MMR08", "MMR09", "MMR10", "MMR11", "MMR12", "MMR13", "MMR14", "MMR15", "MMR16", "MMR17")
```

```{r}
#rename columns 
MMR0517$location_name[MMR0517$location_name == "Iran, Islamic Rep."] <- "Iran, Islamic Republic of"

MMR0517$location_name[MMR0517$location_name == "Korea, Rep."] <- "Korea (Republic of)"

MMR0517$location_name[MMR0517$location_name == "Moldova"] <- "Moldova, Republic of"


gsni_cvd_lifexp_phys_sch_gdp_mmr <- merge(gsni_cvd_lifexp_phys_sch_gdp, MMR0517, by = "location_name", all = TRUE) 
gsni_cvd_lifexp_phys_sch_gdp_mmr <- na.omit(gsni_cvd_lifexp_phys_sch_gdp_mmr, cols = "GSNI_PERIOD") #omit countries without GSNI




```

##COUNTRY INCOME LEVEL

from: https://datahelpdesk.worldbank.org/knowledgebase/articles/906519-world-bank-country-and-lending-groups (historical classification by income)

```{r}
LMH <- read_excel("data/OGHIST.xls", sheet = 3)
LMH <- as.data.table(LMH)

LMH1 <- LMH[5:228, c(2, 25, 30)] #extract required data

LMH2 <- LMH1[-c(2:6), ]

#LMH09 and LMH14 = world bank classification in 2009 & 2014 using GNI per capita 
setnames(LMH2, "...25", "LMH09") 
setnames(LMH2, "...30", "LMH14")
setnames(LMH2,"World Bank Analytical Classifications", "location_name")

LMH2 <- LMH2[-1,]

#rename columns 
LMH2$location_name[LMH2$location_name == "Iran, Islamic Rep."] <- "Iran, Islamic Republic of"
LMH2$location_name[LMH2$location_name == "Korea, Rep."] <- "Korea (Republic of)"
LMH2$location_name[LMH2$location_name == "Moldova"] <- "Moldova, Republic of"

#merge with table
gsni_cvd_lifexp_phys_sch_gdp_mmr_lmh <- merge(gsni_cvd_lifexp_phys_sch_gdp_mmr, LMH2, by = "location_name", all = TRUE) 
gsni_cvd_lifexp_phys_sch_gdp_mmr_lmh <- na.omit(gsni_cvd_lifexp_phys_sch_gdp_mmr_lmh, cols = "GSNI_PERIOD") #omit countries without GSNI

head(gsni_cvd_lifexp_phys_sch_gdp_mmr_lmh)
na.omit(gsni_cvd_lifexp_phys_sch_gdp_mmr_lmh, c("LMH09", "LMH14"), invert = TRUE)


```

```{r}
LMH[c(118, 223, 226, 225),]

#Kyrgzstan, viet nam and yemen are low middle income (names didn't sync)
```



###as.numeric
```{r}
gsni_cvd_lifexp_phys_sch_gdp_mmr_lmh[, twobias := as.numeric(twobias)]
gsni_cvd_lifexp_phys_sch_gdp_mmr_lmh[, twobias := as.numeric(twobias)]
gsni_cvd_lifexp_phys_sch_gdp_mmr_lmh[, onebias := as.numeric(onebias)]
gsni_cvd_lifexp_phys_sch_gdp_mmr_lmh[, nobias := as.numeric(nobias)]
gsni_cvd_lifexp_phys_sch_gdp_mmr_lmh[, political:= as.numeric(political)]
gsni_cvd_lifexp_phys_sch_gdp_mmr_lmh[, economic:= as.numeric(economic)]
gsni_cvd_lifexp_phys_sch_gdp_mmr_lmh[, educational:= as.numeric(educational)]
gsni_cvd_lifexp_phys_sch_gdp_mmr_lmh[, physical := as.numeric(physical)]

gsni_cvd_lifexp_phys_sch_gdp_mmr_lmh[, LEbirth2000 := as.numeric(LEbirth2000)]
gsni_cvd_lifexp_phys_sch_gdp_mmr_lmh[, LEbirth2010 := as.numeric(LEbirth2010)]
gsni_cvd_lifexp_phys_sch_gdp_mmr_lmh[, LEbirth2015 := as.numeric(LEbirth2015)]
gsni_cvd_lifexp_phys_sch_gdp_mmr_lmh[, LEbirth2019 := as.numeric(LEbirth2019)]

gsni_cvd_lifexp_phys_sch_gdp_mmr_lmh[, MLEbirth2000 := as.numeric(MLEbirth2000)]
gsni_cvd_lifexp_phys_sch_gdp_mmr_lmh[, MLEbirth2010 := as.numeric(MLEbirth2010)]
gsni_cvd_lifexp_phys_sch_gdp_mmr_lmh[, MLEbirth2015 := as.numeric(MLEbirth2015)]
gsni_cvd_lifexp_phys_sch_gdp_mmr_lmh[, MLEbirth2019 := as.numeric(MLEbirth2019)]

gsni_cvd_lifexp_phys_sch_gdp_mmr_lmh[, LE602000 := as.numeric(LE602000)]
gsni_cvd_lifexp_phys_sch_gdp_mmr_lmh[, LE602010 := as.numeric(LE602010)]
gsni_cvd_lifexp_phys_sch_gdp_mmr_lmh[, LE602015 := as.numeric(LE602015)]
gsni_cvd_lifexp_phys_sch_gdp_mmr_lmh[, LE602019 := as.numeric(LE602019)]

gsni_cvd_lifexp_phys_sch_gdp_mmr_lmh[, GDP05 := as.numeric(GDP05)]
gsni_cvd_lifexp_phys_sch_gdp_mmr_lmh[, GDP06 := as.numeric(GDP06)]
gsni_cvd_lifexp_phys_sch_gdp_mmr_lmh[, GDP07 := as.numeric(GDP07)]
gsni_cvd_lifexp_phys_sch_gdp_mmr_lmh[, GDP08 := as.numeric(GDP08)]
gsni_cvd_lifexp_phys_sch_gdp_mmr_lmh[, GDP09 := as.numeric(GDP09)]
gsni_cvd_lifexp_phys_sch_gdp_mmr_lmh[, GDP10 := as.numeric(GDP10)]
gsni_cvd_lifexp_phys_sch_gdp_mmr_lmh[, GDP11 := as.numeric(GDP11)]
gsni_cvd_lifexp_phys_sch_gdp_mmr_lmh[, GDP12 := as.numeric(GDP12)]
gsni_cvd_lifexp_phys_sch_gdp_mmr_lmh[, GDP13 := as.numeric(GDP13)]
gsni_cvd_lifexp_phys_sch_gdp_mmr_lmh[, GDP14 := as.numeric(GDP14)]
gsni_cvd_lifexp_phys_sch_gdp_mmr_lmh[, GDP15 := as.numeric(GDP15)]
gsni_cvd_lifexp_phys_sch_gdp_mmr_lmh[, GDP16 := as.numeric(GDP16)]
gsni_cvd_lifexp_phys_sch_gdp_mmr_lmh[, GDP17 := as.numeric(GDP17)]

gsni_cvd_lifexp_phys_sch_gdp_mmr_lmh[, MMR05 := as.numeric(MMR05)]
gsni_cvd_lifexp_phys_sch_gdp_mmr_lmh[, MMR06 := as.numeric(MMR06)]
gsni_cvd_lifexp_phys_sch_gdp_mmr_lmh[, MMR07 := as.numeric(MMR07)]
gsni_cvd_lifexp_phys_sch_gdp_mmr_lmh[, MMR08 := as.numeric(MMR08)]
gsni_cvd_lifexp_phys_sch_gdp_mmr_lmh[, MMR09 := as.numeric(MMR09)]
gsni_cvd_lifexp_phys_sch_gdp_mmr_lmh[, MMR10 := as.numeric(MMR10)]
gsni_cvd_lifexp_phys_sch_gdp_mmr_lmh[, MMR11 := as.numeric(MMR11)]
gsni_cvd_lifexp_phys_sch_gdp_mmr_lmh[, MMR12 := as.numeric(MMR12)]
gsni_cvd_lifexp_phys_sch_gdp_mmr_lmh[, MMR13 := as.numeric(MMR13)]
gsni_cvd_lifexp_phys_sch_gdp_mmr_lmh[, MMR14 := as.numeric(MMR14)]
gsni_cvd_lifexp_phys_sch_gdp_mmr_lmh[, MMR15 := as.numeric(MMR15)]
gsni_cvd_lifexp_phys_sch_gdp_mmr_lmh[, MMR16 := as.numeric(MMR16)]
gsni_cvd_lifexp_phys_sch_gdp_mmr_lmh[, MMR17 := as.numeric(MMR17)]

gsni_cvd_lifexp_phys_sch_gdp_mmr_lmh[, GSNI_PERIOD := as.factor(GSNI_PERIOD)]

gsni_cvd_lifexp_phys_sch_gdp_mmr_lmh[, LMH09 := factor(LMH09, 
                           levels = c("L", "LM", "UM", "H"),
                           labels = c("low income", "lower middle income", "upper middle income", "high income")
                      )]
gsni_cvd_lifexp_phys_sch_gdp_mmr_lmh[, LMH14 := factor(LMH14, 
                           levels = c("L", "LM", "UM", "H"),
                           labels = c("low income", "lower middle income", "upper middle income", "high income")
                      )]

```


```{r}

glimpse(gsni_cvd_lifexp_phys_sch_gdp_mmr_lmh)
```

Create columns - ratio of female to male CVD for CVD14 and CVD17 

```{r}

gsni_cvd_lifexp_phys_sch_gdp_mmr_lmh[, `:=` (cvdratio14 = cvd14fem / cvd14male,
                                             cvdratio17 = cvd17fem /cvd17male)]

```

##examine missing values and insert replacements
 
###2014
```{r}

#examine NAs for 2014
edit <- cbind(gsni_cvd_lifexp_phys_sch_gdp_mmr_lmh) #make new table for editing (cbind) means a duplicate will be made otherwise changing 'edit' will also change the gsni table


na.omit(edit, c("cvd14fem", "cvd14male", "cvd14both", "LEbirth2015", "MLEbirth2015", "cvdratio14",  "phy14", "scl14", "GDP14", "MMR14"), invert = TRUE) #28 rows with one missing value for 2014


edit
```

####cvd14fem
```{r}
na.omit(edit, c("cvd14fem"), invert = TRUE) #missing cvdfem -> no missing

```

####cvd14male
```{r}
na.omit(edit, c("cvd14male"), invert=TRUE) #no missing
```

####cvd14
```{r}
na.omit(edit, c("cvd14both"), invert = TRUE) #missing cvdboth -> no missing

```

####cvdratio14
```{r}
na.omit(edit, c("cvdratio14"), invert = TRUE) #missing cvdboth -> no missing

```


####LEbirth2015
```{r}
na.omit(edit, c("LEbirth2015"), invert = TRUE) #missing LE at birth (Andorra, Palestine and UK)
```

```{r}

edit[c("Andorra", "Palestine, State of", "United Kingdom"), c("location_name", "LEbirth2000", "LEbirth2010", "LEbirth2015", "LEbirth2019")]

#no data points that can be used to fill LE at birth for these countries
```
Have data for UK from IHME from 2015 https://vizhub.healthdata.org/gbd-compare/le
LE at birth for females 2015 = 84.8

```{r}

edit$LEbirth2015[68] <- 82.83 #inserted UK value

edit[68,LEbirth2015]

```

Have data for Andorra from IHME from 2015 https://vizhub.healthdata.org/gbd-compare/le
LE at birth for females 2015 = 84.8

```{r}
edit$LEbirth2015[2] <- 84.8 #inserted Andorra value

edit[2,LEbirth2015]
```

Have data for Palestine from IHME from 2015 https://vizhub.healthdata.org/gbd-compare/le
LE at birth for females 2015 = 76.22
```{r}
edit$LEbirth2015[48] <- 76.22 #inserted Palestine value

edit[48,LEbirth2015]
```

####MLE2015
```{r}
na.omit(edit, c("MLEbirth2015"), invert = TRUE) #missing LE at birth (Andorra, Palestine and UK)


edit[c("Andorra", "Palestine, State of", "United Kingdom"), c("location_name", "MLEbirth2000", "MLEbirth2010", "MLEbirth2015", "MLEbirth2019")] 
```

Have data for UK from IHME from 2015 https://vizhub.healthdata.org/gbd-compare/le
LE at birth for males 2015 in UK = 79.22

```{r}

edit$MLEbirth2015[68] <- 79.22 #inserted UK value

edit[68,MLEbirth2015]

```

Have data for Andorra from IHME from 2015 https://vizhub.healthdata.org/gbd-compare/le
LE at birth for males 2015 = 79.56

```{r}
edit$MLEbirth2015[2] <- 79.56 #inserted Andorra value

edit[2,MLEbirth2015]
```

Have data for Palestine from IHME from 2015 https://vizhub.healthdata.org/gbd-compare/le
LE at birth for males 2015 = 72.46
```{r}
edit$MLEbirth2015[48] <- 72.46 #inserted Palestine value

edit[48,MLEbirth2015]
```


####phy14
```{r}
na.omit(edit, c("phy14"), invert = TRUE) #missing phy (23 rows)
```
```{r}

naphy <- edit[, c("location_name", "phy09","phy10", "phy11", "phy12", "phy13", "phy14", "phy15", "phy16", "phy17")]
naphy #physican columns 09-17

na.omit(naphy, c("phy14"), invert = TRUE) #able to see if can replace values


```

REPLACE

Algeria	-2016
Andorra		- 2015
Argentina	- 2013
Brazil	- 2013
Ecuador	- 2015
Ethiopia	- 2016
Ghana	- 2013
Haiti - 2015
Indonesia	- 2015
Kyrgyzsta - NA
Malaysia	- 2015
Mali	- 2016
Nigeria		- 2013
Palestine, State of		-NA
Peru - 2016		
Philippines - 2017
Romania	- 2013
Thailand	- 2015	
Trinidad and Tobago	- 2015
Uruguay - 2016
Viet Nam	- NA
Yemen	- NA
Zambia	- 2016
```{r}
#need to create table with missing phy14 values in one column the can use rows_patch to insert new values

Algeriaphy <- edit["Algeria", c("location_name", "phy16")]
Andorraphy <- edit["Andorra", c("location_name", "phy15")]
Argentinaphy <- edit["Argentina", c("location_name", "phy13")]
Brazilphy <- edit["Brazil", c("location_name", "phy13")]
Ecuadorphy <- edit["Ecuador", c("location_name", "phy15")]
Ethiopiaphy	<- edit["Ethiopia", c("location_name", "phy16")]
Ghanaphy <- edit["Ghana", c("location_name", "phy13")]	
Haitiphy <- edit["Haiti", c("location_name", "phy15")]
Indonesiaphy	<- edit["Indonesia", c("location_name", "phy15")]
#Kyrgyzsta - NA
Malaysiaphy	<- edit["Malaysia", c("location_name", "phy15")]
Maliphy	<- edit["Mali", c("location_name", "phy16")]
Nigeriaphy		<- edit["Nigeria", c("location_name", "phy13")]
#Palestine, State of		-NA
Peruphy <- edit["Peru", c("location_name", "phy16")]	
Philippinesphy <- edit["Philippines", c("location_name", "phy17")]
Romaniaphy	<- edit["Romania", c("location_name", "phy13")]
Thailandphy	<- edit["Thailand", c("location_name", "phy15")]
TrinidadandTobagophy	<- edit["Trinidad and Tobago", c("location_name", "phy15")]
Uruguayphy <- edit["Uruguay", c("location_name", "phy16")]	
#Viet Nam	- NA
#Yemen	- NA
Zambiaphy <-	edit["Zambia", c("location_name", "phy16")]	


phy14edit <- rbind(Algeriaphy, Andorraphy, Argentinaphy, Brazilphy, Ecuadorphy, Ethiopiaphy, Ghanaphy, Haitiphy, Indonesiaphy, Malaysiaphy, Maliphy, Nigeriaphy, Peruphy, Philippinesphy, Romaniaphy, Thailandphy, TrinidadandTobagophy, Uruguayphy,Zambiaphy,   use.names=FALSE)
setnames(phy14edit, "phy16", "phy14edit") #new table with values to insert into the phy14 column
```

```{r}
setnames(edit, "phy14", "phy14edit")

edit_phy14 <- rows_patch(edit, phy14edit)
edit_phy14[,c("location_name", "phy14edit")]
na.omit(edit_phy14, "phy14edit", invert=TRUE)

#table "edit_phy14" now has column "phy14edit" which has inserted values from other years
# 4 NAs in total for phy
```





####scl14

```{r}
na.omit(edit_phy14, c("scl14"), invert = TRUE) # 6 missing scl 2014

nascl <- edit_phy14[, c("location_name", "scl05", "scl06", "scl07", "scl08", "scl09","scl10", "scl11", "scl12", "scl13", "scl14", "scl15", "scl16", "scl17")]
nascl #scl columns 09-17

na.omit(nascl, c("scl14"), invert = TRUE) #able to see if can replace values, can't replace with any , still 6 NAs

```
SCL no values
Iran, Islamic Republic of	NA	
Korea (Republic of)	NA	
Moldova, Republic of	NA		
Palestine, State of	NA
Russian Federation NA
Viet Nam NA


####GDP14
```{r}
na.omit(edit_phy14, c("GDP14"), invert = TRUE) # 7 missing GDP2014

naGDP <- edit_phy14[, c("location_name", "GDP05", "GDP06", "GDP07", "GDP08", "GDP09","GDP10", "GDP11", "GDP12", "GDP13", "GDP14", "GDP15", "GDP16", "GDP17")]
naGDP #GDP columns 09-17

na.omit(naGDP, c("GDP14"), invert = TRUE) #able to see if can replace values, can't replace with any , still 7 NAs

```
GDP (no values)
Iran, Islamic Republic of	NA	
Korea (Republic of)	NA	
Kyrgyzstan	NA	
Moldova, Republic of	NA	
Palestine, State of	NA	
Viet Nam	NA		
Yemen NA

####MMR14
```{r}
na.omit(edit_phy14, c("MMR14"), invert = TRUE) # 5 missing MMR2014

naMMR <- edit_phy14[, c("location_name", "MMR05", "MMR06", "MMR07", "MMR08", "MMR09","MMR10", "MMR11", "MMR12", "MMR13", "MMR14", "MMR15", "MMR16", "MMR17")]
naMMR #MMR columns 09-17

na.omit(naMMR, c("MMR14"), invert = TRUE) #able to see if can replace values, can't replace any

```
MMR NAs (no values)
Andorra	NA	
Kyrgyzstan	NA	
Palestine, State of	NA
Viet Nam	NA
Yemen NA


####final14
```{r}
analysisratio14 <- edit_phy14[, c("location_name", "GSNI_PERIOD", "twobias","cvd14fem", "cvd14male", "cvd14both", "cvdratio14" , "LEbirth2015", "MLEbirth2015", "phy14edit", "scl14", "GDP14", "MMR14")]
glimpse(analysisratio14)
# table name = analysis14
```
```{r}
na.omit(analysisratio14, c("cvd14fem", "cvd14male", "cvd14both", "cvdratio14" , "LEbirth2015", "MLEbirth2015", "phy14edit", "scl14", "GDP14", "MMR14"), invert = TRUE)

#now have 9 rows with missing data 
```


 
###2017
```{r}

#examine NAs for 2017
edit17 <- cbind(gsni_cvd_lifexp_phys_sch_gdp_mmr_lmh)

na.omit(edit17, c("cvd17fem","cvd14male","cvdratio17", "cvd17both", "LEbirth2019", "MLEbirth2019", "phy17", "scl17", "GDP17", "MMR17"), invert = TRUE) #30 rows with one missing value for 2017

```

####cvd17fem
```{r}
na.omit(edit17, c("cvd17fem"), invert = TRUE) #missing cvdfem -> no missing

```

####cvd14male
```{r}
na.omit(edit17, c("cvd17male"), invert = TRUE) #missing cvdmale -> no missing

```

####cvdratio17
```{r}
na.omit(edit17, c("cvdratio17"), invert = TRUE) #-> no missing

```

####cvd17
```{r}
na.omit(edit17, c("cvd17both"), invert = TRUE) #missing cvdboth -> no missing

```
####LEbirth2019
```{r}
na.omit(edit17, c("LEbirth2015"), invert = TRUE) #missing LE at birth (Andorra, Palestine and UK)
```

```{r}

edit17[c("Andorra", "Palestine, State of", "United Kingdom"), c("location_name", "LEbirth2000", "LEbirth2010", "LEbirth2015", "LEbirth2019")]

#no data points that can be used to fill LE at birth for these countries
```
Can add in data from IMHE https://vizhub.healthdata.org/gbd-compare/le
LE at birth for females 2019 
UK[68] = 82.88
Andorra[2]= 84.95
Palestine[48] = 76.47


```{r}
edit17$LEbirth2019[68] <- 82.88
edit17$LEbirth2019[2] <- 84.95
edit17$LEbirth2019[48] <- 76.47

edit17[c("Andorra", "Palestine, State of", "United Kingdom"), c("location_name", "LEbirth2000", "LEbirth2010", "LEbirth2015", "LEbirth2019")]

```

####MLEbirth2019
```{r}
na.omit(edit17, c("MLEbirth2015"), invert = TRUE) #missing LE at birth (Andorra, Palestine and UK)
```

```{r}

edit17[c("Andorra", "Palestine, State of", "United Kingdom"), c("location_name", "MLEbirth2000", "MLEbirth2010", "MLEbirth2015", "MLEbirth2019")]

#no data points that can be used to fill LE at birth for these countries
```
Can add in data from IMHE https://vizhub.healthdata.org/gbd-compare/le
LE at birth for males 2019 
UK[68] = 79.24
Andorra[2]= 79.8
Palestine[48] = 73.19


```{r}
edit17$MLEbirth2019[68] <- 79.24
edit17$MLEbirth2019[2] <- 79.8
edit17$MLEbirth2019[48] <- 73.19

edit17[c("Andorra", "Palestine, State of", "United Kingdom"), c("location_name", "MLEbirth2000", "MLEbirth2010", "MLEbirth2015", "MLEbirth2019")]

```


####phy17
```{r}
na.omit(edit17, c("phy17"), invert = TRUE) #missing phy (26 rows)
```



```{r}

naphy17 <- edit17[, c("location_name", "phy09","phy10", "phy11", "phy12", "phy13", "phy14", "phy15", "phy16", "phy17")]
naphy17 #physican columns 09-17

na.omit(naphy17, c("phy17"), invert = TRUE) #able to see if can replace values


```

REPLACE

```{r}
Andorra	<- edit17["Andorra", c("location_name", "phy15")]
Azerbaijan	<- edit17["Azerbaijan", c("location_name", "phy14")]
Belarus	<- edit17["Belarus", c("location_name", "phy15")]
Bulgaria	<- 	edit17["Bulgaria", c("location_name", "phy15")]
Cyprus	<- edit17["Cyprus", c("location_name", "phy16")]
Ecuador <- edit17["Ecuador", c("location_name", "phy16")]
Finland		<- edit17["Finland", c("location_name", "phy16")]
Haiti		<-	edit17["Haiti", c("location_name", "phy15")]
Japan		<- edit17["Japan", c("location_name", "phy16")]
Kazakhstan	<- edit17["Kazakhstan", c("location_name", "phy14")]
Kuwait	<-	edit17["Kuwait", c("location_name", "phy15")]
#Kyrgyzstan		<- NA
Malaysia		<- edit17["Malaysia", c("location_name", "phy15")]
Mali		<- edit17["Mali", c("location_name", "phy16")]
Nigeria		<- edit17["Nigeria", c("location_name", "phy16")]
#Palestine, State of			<- NA
Peru		<- edit17["Peru", c("location_name", "phy16")]
RussianFederation	<- edit17["Russian Federation", c("location_name", "phy16")]
Serbia		<- edit17["Serbia", c("location_name", "phy16")]
Singapore	<- edit17["Singapore", c("location_name", "phy16")]
Sweden		<- edit17["Sweden", c("location_name", "phy16")]
Ukraine		<- edit17["Ukraine", c("location_name", "phy14")]
Uzbekistan		<- edit17["Uzbekistan", c("location_name", "phy14")]
#Viet Nam	<- NA
#Yemen		<- NA
Zambia	<- edit17["Zambia", c("location_name", "phy16")]
```

need to create table with missing phy14 values in one column the can use rows_patch to insert new values

```{r}
phy17edit <- rbind(Andorra,
Azerbaijan,
Belarus	,
Bulgaria	,
Cyprus	,
Ecuador ,
Finland		,
Haiti	,
Japan		,
Kazakhstan	,
Kuwait	,
Malaysia	,
Mali	,
Nigeria		,
Peru		,
RussianFederation,
Serbia	,
Singapore	,
Sweden	,
Ukraine	,
Uzbekistan	,
Zambia	,
use.names=FALSE)

setnames(phy17edit, "phy15", "phy17edit") #new table with values to insert into the phy14 column
phy17edit
```



```{r}
setnames(edit17, "phy17", "phy17edit")

edit_phy17 <- rows_patch(edit17, phy17edit)
edit_phy17[,c("location_name", "phy17edit")]
na.omit(edit_phy17, "phy17edit", invert=TRUE)

#table "edit_phy17" now has column "phy17edit" which has inserted values from other years
# 4 NAs in total for phy

```




####scl17

```{r}
na.omit(edit_phy17, c("scl17"), invert = TRUE) # 6 missing scl 2017

nascl <- edit_phy17[, c("location_name", "scl05", "scl06", "scl07", "scl08", "scl09","scl10", "scl11", "scl12", "scl13", "scl14", "scl15", "scl16", "scl17")]
nascl #scl columns 09-17

na.omit(nascl, c("scl17"), invert = TRUE) #able to see if can replace values, can't replace with any , still 6 NAs


```
SCL no values
Iran, Islamic Republic of	NA	
Korea (Republic of)	NA	
Moldova, Republic of	NA		
Palestine, State of	NA
Russian Federation NA
Viet Nam NA


####GDP17
```{r}
na.omit(edit_phy17, c("GDP17"), invert = TRUE) # 7 missing GDP2017

naGDP <- edit_phy17[, c("location_name", "GDP05", "GDP06", "GDP07", "GDP08", "GDP09","GDP10", "GDP11", "GDP12", "GDP13", "GDP14", "GDP15", "GDP16", "GDP17")]
naGDP #GDP columns 09-17

na.omit(naGDP, c("GDP17"), invert = TRUE) #able to see if can replace values, can't replace with any , still 7 NAs

```
GDP (no values)
Iran, Islamic Republic of	NA	
Korea (Republic of)	NA	
Kyrgyzstan	NA	
Moldova, Republic of	NA	
Palestine, State of	NA	
Viet Nam	NA		
Yemen NA

####MMR17
```{r}
na.omit(edit_phy17, c("MMR17"), invert = TRUE) # 5 missing MMR2017

naMMR <- edit_phy17[, c("location_name", "MMR05", "MMR06", "MMR07", "MMR08", "MMR09","MMR10", "MMR11", "MMR12", "MMR13", "MMR14", "MMR15", "MMR16", "MMR17")]
naMMR #MMR columns 09-17

na.omit(naMMR, c("MMR17"), invert = TRUE) #able to see if can replace values, can't replace any

```
MMR NAs (no values)
Andorra	NA	
Kyrgyzstan	NA	
Palestine, State of	NA
Viet Nam	NA
Yemen NA


####final17
```{r}

analysisratio17 <- edit_phy17[, c("location_name", "GSNI_PERIOD", "twobias", "cvd17fem", "cvd17male", "cvdratio17", "cvd17both", "LEbirth2019", "MLEbirth2019", "phy17edit", "scl17", "GDP17", "MMR17")]
glimpse(analysisratio17)
# table name = analysisratio17


na.omit(edit_phy17, c("twobias", "cvd17fem", "cvd17male", "cvdratio17", "cvd17both", "LEbirth2019", "MLEbirth2019", "phy17edit", "scl17", "GDP17", "MMR17"), invert = TRUE) 

#now have 9 rows with one missing 

```




```{r}
write.csv(analysisratio14,"analysisratio14.csv")
write.csv(analysisratio17,"analysisratio17.csv")
```



```{r}

countrylist <- cbind(edit[, c("location_name", "LMH14")])
countrylist
write.csv(countrylist, "countrylist.csv")

countrylist75 <- cbind(gsni_cvd_lifexp_phys_sch_gdp_mmr_lmh[, c("location_name", "LMH14")])
countrylist75
write.csv(countrylist75, "countrylist75.csv")
```


