---
title: "Notes"
author: "Iona Lyell"
date: "27/01/2022"
output: html_document
---

To do:


letter to editor: updated

Abstract: updated
 
Intro: updated

Methods: updated

Results: 
updated

Disucssion: 

unmeasured confounding examples: e.g. diet, parity

Appendix: countries included in each analysis, *review appendix and tables*

Share results plus specific Qs for reviewers, 
- what more to add re gender...struggling with this
- thought on interpretation


Plan to submit to BMJ mid-Feb (secondary journal options Circulation and JACC)


Regression table 1: female CVD mortality , male CVD mortality
```{r, echo = FALSE, results='asis'}
stargazer(mf1, mf2, mf3, mf4, type="html",
out="man_results_regressiontable_1_n.doc",
intercept.bottom = F, intercept.top = T,
digits=2,
model.names = F,
single.row = T,
star.char = c("*", "**", "***"),
          star.cutoffs = c(.05, .01, .001),
notes = c("* p<0.05; ** p<0.01; *** p<0.001"),
notes.append=FALSE,
title = "Results table 1 n",
covariate.labels = c("constant", "GSNI2","physicians per 1000", "mean years of schooling", "GDP per capita", "GSNI data collection period 2010-2014"),
dep.var.labels = c("female CVD mortality", "male CVD mortality"),
omit.stat = c("ser")
                               )
```
Regression table 2: CVD mortality , ratio CVD mortality (Qatar excluded)
```{r, echo = FALSE, results='asis'}
stargazer(mf5, mf6, mf7, mf8, type="html",
out="man_results_regressiontable_2_n.doc",
intercept.bottom = F, intercept.top = T,
digits=2,
model.names = F,
single.row = T,
star.char = c("*", "**", "***"),
          star.cutoffs = c(.05, .01, .001),
notes = c("* p<0.05; ** p<0.01; *** p<0.001"),
notes.append=FALSE,
title = "Results table 2",
covariate.labels = c("constant", "GSNI2","physicians per 1000", "mean years of schooling", "GDP per capita", "GSNI data collection period 2010-2014"),
dep.var.labels = c("CVD mortality", "female to male CVD mortality ratio"),
omit.stat = c("ser")
                               )
```
Regression table 3: Female and male life expectancy
```{r}


stargazer(mf9, mf10, mf11, mf12, type="html",
out="man_results_regressiontable_3_n.doc",
intercept.bottom = F, intercept.top = T,
digits=2,
model.names = F,
single.row = T,
star.char = c("*", "**", "***"),
          star.cutoffs = c(.05, .01, .001),
notes = c("* p<0.05; ** p<0.01; *** p<0.001"),
notes.append=FALSE,
title = "Results table 3",
dep.var.labels = c("female life expectancy", "male life expectancy"),
covariate.labels = c("constant", "GSNI","log (physicians per 1000)", "mean years of schooling", "log (GDP per capita)", "GSNI data collection period 2010-2014", "MMR"),
omit.stat = c("ser")
                               )


```
 Regression table 4: sensitivityfemale CVD mortality , male CVD mortality
```{r, echo = FALSE, results='asis'}
stargazer(mf13, mf14, mf15, mf16, type="html",
out="man_results_regressiontable_4_n.doc",
intercept.bottom = F, intercept.top = T,
digits=2,
model.names = F,
single.row = T,
star.char = c("*", "**", "***"),
          star.cutoffs = c(.05, .01, .001),
notes = c("* p<0.05; ** p<0.01; *** p<0.001"),
notes.append=FALSE,
title = "Results table 4",
covariate.labels = c("constant", "GSNI2","physicians per 1000", "mean years of schooling", "GDP per capita", "GSNI data collection period 2010-2014"),
dep.var.labels = c("female CVD mortality 17", "male CVD mortality 17"),
omit.stat = c("ser")
                               )
```



Regression table 5: sensitivity CVD mortality , ratio CVD mortality (Qatar excluded)
```{r, echo = FALSE, results='asis'}
stargazer(mf17, mf18, mf19, mf20, type="html",
out="man_results_regressiontable_5_n.doc",
intercept.bottom = F, intercept.top = T,
digits=2,
model.names = F,
single.row = T,
star.char = c("*", "**", "***"),
          star.cutoffs = c(.05, .01, .001),
notes = c("* p<0.05; ** p<0.01; *** p<0.001"),
notes.append=FALSE,
title = "Results table 5",
covariate.labels = c("constant", "GSNI2","physicians per 1000", "mean years of schooling", "GDP per capita", "GSNI data collection period 2010-2014"),
dep.var.labels = c("CVD mortality 17", "female to male CVD mortality ratio 17"),
omit.stat = c("ser")
                               )
```




Regression table 6: sensitivity Female and male life expectancy
```{r}


stargazer(mf21, mf22, mf23, mf24, type="html",
out="man_results_regressiontable_6_n.doc",
intercept.bottom = F, intercept.top = T,
digits=2,
model.names = F,
single.row = T,
star.char = c("*", "**", "***"),
          star.cutoffs = c(.05, .01, .001),
notes = c("* p<0.05; ** p<0.01; *** p<0.001"),
notes.append=FALSE,
title = "Results table 6",
dep.var.labels = c("female life expectancy 19", "male life expectancy 19"),
covariate.labels = c("constant", "GSNI","log (physicians per 1000)", "mean years of schooling", "log (GDP per capita)", "GSNI data collection period 2010-2014", "MMR"),
omit.stat = c("ser")
                               )


```

