---
title: "dissertation analysis 2014"
output: github_document
editor_options: 
  chunk_output_type: inline
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

###load packages
```{r}


library(qwraps2)
options(qwraps2_markup = "markdown")
library(readr)
library(readxl)
library(data.table)
library(ggplot2)
library(dplyr)
library(skimr)
library(tidyverse)
library(broom)
library(SmartEDA)


analysis14 <- read_csv("analysisratio14.csv")
setDT(analysis14)

analysis14[, GSNI_PERIOD := as.factor(GSNI_PERIOD)]
analysis14[, "...1":= NULL]


```






#2014

# CODE BOOK (analysis14)

+-------------------------+----------------------------------------------------+
| Variable                | Description                                        |
+=========================+====================================================+
| location_name           | country                                            |
+-------------------------+----------------------------------------------------+
| GSNI_PERIOD             | year data collection took place for GSNI survey    |
|                         |  2005-2009 or 2010-2014                            |
+-------------------------+----------------------------------------------------+
| twobias                 | share of people with at least two bias             |
+-------------------------+----------------------------------------------------+
| cvd14fem                | age standardised CVD mortality rates for women per |            |                         |    100,000 population                              |
|                         |    (from global burden of disease)                 |            
|                         |     2014                                           |
+-------------------------+----------------------------------------------------+
| cvd14male               | as above for whole males                           |
+-------------------------+----------------------------------------------------+
| cvd14both               | as above for whole population                      |
+-------------------------+----------------------------------------------------+
| cvdratio14              | ratio female/male CVD mortality rates              |
+-------------------------+----------------------------------------------------+
| LEbirth2015             |  life expectancy at birth for                      |         
|                         |  females in 2015, from WHO (years)                 |
+-------------------------+----------------------------------------------------+
| MLEbirth2015            |  life expectancy at birth for                      |         
|                         |  males in 2015, from WHO (years)                   |
+-------------------------+----------------------------------------------------+
| phy14edit               | physicians per 1000 population 2014 (world bank)   |
|                         |  data added from other years n = 19 (2013-17)      |
+-------------------------+----------------------------------------------------+
| scl14                   | mean years of school 2014 (ourworldindata)         |
+-------------------------+----------------------------------------------------+
| GDP14                   | Gross domestic product per capita (current US      |            |                         | dollar)  2014  (world bank)                        |
+-------------------------+----------------------------------------------------+
| MMR14                   | Maternal Mortality Ratio 2014   (world bank)       |
+-------------------------+----------------------------------------------------+


#Descriptive analysis
```{r}
glimpse(analysis14)
summary(analysis14)
```



```{r}
analysis14copy <- copy(analysis14)
analysis14copy2 <- copy(analysis14)

#create data table without MMR
analysis14noMMR <- cbind(analysis14copy[, MMR14:=NULL])

#create table without MMR and without NAs
analysis14noMMRnoNA <- cbind(na.exclude(analysis14noMMR))

#creat table with MMR with no NAs
analysis14noNA <- cbind(na.exclude(analysis14copy2)) #exclude rows with NA as will prevent model comparison

#just use analysis14 for all analysis as will exclude NAs
```

list of countries included
```{r}
analysis14noNA[,location_name] #66 rows (66 data points for MMR analysis)
analysis14noMMRnoNA[,location_name] #67 rows 
analysis14[,location_name] #75 rows
```




#####Histograms
```{r}
#ggplot(analysis14noNA, aes(x=twobias))+geom_histogram(binwidth=5, color = "Black", fill = "Pink")

analysis14noNA[,hist(twobias)]
analysis14noNA[,hist(cvd14fem)]
analysis14noNA[,hist(cvd14male)]
analysis14noNA[,hist(cvdratio14)]
analysis14noNA[,hist(cvd14both)]
analysis14noNA[,hist(LEbirth2015)]
analysis14noNA[,hist(MLEbirth2015)]
analysis14noNA[,hist(phy14edit)]
analysis14noNA[,hist(scl14)]
analysis14noNA[,hist(GDP14)]
analysis14noNA[,hist(MMR14)]

#no continuous variables variables normally distributed

summary(analysis14noNA
        )
```

#####summary tables
https://statsandr.com/blog/descriptive-statistics-in-r/
https://cran.r-project.org/web/packages/qwraps2/vignettes/summary-statistics.html

```{r}
our_summary1 <-
  list("GSNI index 2 or more bias" =
       list("min"       = ~ round(min(twobias),1),
            "max"       = ~ round(max(twobias),1),
            "median (IQR)" = ~ median_iqr(twobias),
        "missing" = ~sum(is.na(twobias))),
       "Female CVD age adjusted mortality rate (per 100 000 population)" =
       list("min"       = ~ round(min(cvd14fem),1),
            "max"       = ~ round(max(cvd14fem),1),
            "median (IQR)" = ~ median_iqr(cvd14fem),
             "missing" = ~sum(is.na(cvd14fem))),
       "Male CVD age adjusted mortality rate (per 100 000 population)" =
       list("min"       = ~ round(min(cvd14male),1),
            "max"       = ~ round(max(cvd14male),1),
            "median (IQR)" = ~ median_iqr(cvd14male),
             "missing" = ~sum(is.na(cvd14male))),
       "Population CVD age adjusted mortality rate (per 100 000 population)" =
       list("min"       = ~ round(min(cvd14both),1),
            "max"       = ~ round(max(cvd14both),1),
            "median (IQR)" = ~ median_iqr(cvd14both),
             "missing" = ~sum(is.na(cvd14both))),
       "Ratio of female to male CVD mortality rates" =
       list("min"       = ~ round(min(cvdratio14),1),
            "max"       = ~ round(max(cvdratio14),1),
            "median (IQR)" = ~ median_iqr(cvdratio14),
             "missing" = ~sum(is.na(cvdratio14))),
       "Female life expectancy at birth" =
       list("min"       = ~ round(min(LEbirth2015),1),
            "max"       = ~ round(max(LEbirth2015),1),
            "median (IQR)" = ~ median_iqr(LEbirth2015),
             "missing" = ~sum(is.na(LEbirth2015))),
        "Male life expectancy at birth" =
       list("min"       = ~ round(min(MLEbirth2015),1),
            "max"       = ~ round(max(MLEbirth2015),1),
            "median (IQR)" = ~ median_iqr(MLEbirth2015),
             "missing" = ~sum(is.na(MLEbirth2015))),
       "Physicians per 1000 population" =
       list("min"       = ~ round(min(phy14edit, na.rm= TRUE),1), 
            "max"       = ~ round(max(phy14edit, na.rm= TRUE),1), 
            "median (IQR)" = ~ median_iqr(phy14edit, na_rm = TRUE),
             "missing" = ~sum(is.na(phy14edit))),
       "GDP per capita" =
       list("min"       = ~ round(min(GDP14, na.rm= TRUE),1),
            "max"       = ~ round(max(GDP14, na.rm= TRUE),1),
            "median (IQR)" = ~ median_iqr(GDP14, na_rm = TRUE),
             "missing" = ~sum(is.na(GDP14))),
        "Maternal Mortality Ratio" =
       list("min"       = ~ round(min(MMR14, na.rm= TRUE),1),
            "max"       = ~ round(max(MMR14, na.rm= TRUE),1),
            "median (IQR)" = ~ median_iqr(MMR14, na_rm = TRUE),
             "missing" = ~sum(is.na(MMR14))),
       "Mean years of schooling" =
       list("min"       = ~ round(min(scl14, na.rm= TRUE),1),
            "max"       = ~ round(max(scl14, na.rm= TRUE),1),
            "median (IQR)" = ~ median_iqr(scl14, na_rm = TRUE),
             "missing" = ~sum(is.na(scl14))),
       "Year of Index Collection" =
       list("2005-2009" = ~ n_perc0(GSNI_PERIOD == levels(analysis14$GSNI_PERIOD)[1]),
            "2010-2014"  = ~ n_perc0(GSNI_PERIOD == levels(analysis14$GSNI_PERIOD)[2])))

            
```






```{r}

whole14 <- summary_table(analysis14, our_summary1)
print(whole14,
      rtitle = "Summary Statistics",
      cnames = paste("n = ", nrow(results14)))
```



#OUTCOME 1 - CVD mortality female

##Scatter plots

###CVD mort (female) ~ GSNI

```{r}
ggplot(analysis14, aes(x = twobias, y = cvd14fem)) +
  geom_point(color = "violetred3")+
  geom_smooth(method = "lm")+
         labs(x = "percentage with 2 GSNI bias (%)",
                y = "female age standardised CVD mortaltiy rate (per 100k)",
                title = "Scatter plot of GSNI (2 bias) and female CVD mortality")

analysis14noNA[,cor.test(twobias,cvd14fem,method = "spearman")]

```


###CVD mort (female) ~ physicians per 1000

```{r}
ggplot(analysis14, aes(x = phy14edit, y = cvd14fem)) +
  geom_point(color = "violetred3")+
  geom_smooth(method = "lm")+
         labs(x = "physicians per 1000 population",
                y = "female age standardised CVD mortaltiy rate (per 100k)",
                title = "Scatter plot of phycisians per 1000 and female CVD mortality")

analysis14noNA[,cor.test(cvd14fem,phy14edit,method = "spearman")]

```

###CVD mort (female) ~ mean years of schooling

```{r}
ggplot(analysis14, aes(x = scl14, y = cvd14fem)) +
  geom_point(color = "violetred3")+
  geom_smooth(method = "lm")+
         labs(x = "mean years of schooling",
                y = "female age standardised CVD mortaltiy rate (per 100k)",
                title = "Mean years of schooling and female CVD mortality")

analysis14noNA[,cor.test(scl14,cvd14fem,method = "spearman")]

```

###CVD mort (female) ~ GDP per capita

```{r}
ggplot(analysis14, aes(x =GDP14, y = cvd14fem)) +
  geom_point(color = "violetred3")+
  geom_smooth(method = "lm")+
         labs(x = "GDP per capita",
                y = "female age standardised CVD mortaltiy rate (per 100k)",
                title = "GDP per capita and female CVD mortality")

analysis14noNA[,cor.test(GDP14,cvd14fem,method = "spearman")]
```

###CVD mort (female) ~ MMR
```{r}
ggplot(analysis14, aes(x = MMR14, y = cvd14fem)) +
  geom_point(color = "violetred3")+
  geom_smooth(method = "lm")+
         labs(x = "Maternal Mortality Ratio",
                y = "female age standardised CVD mortaltiy rate (per 100k)",
                title = "MMR and female CVD mortality")

analysis14noNA[,cor.test(MMR14,cvd14fem,method = "spearman")]

```

###CVD mort (female) ~ year of index
```{r}

outlier14 <- analysis14[cvd14fem>500,] #set up to label extreme values
  
#boxplot separated by period of collection
ggplot(analysis14, aes(x = GSNI_PERIOD, y = cvd14fem)) +
  geom_boxplot(color = "violetred3", notch= TRUE) +
         labs(x = "GSNI period",
                y = "female age standardised CVD mortaltiy rate",
                title = "GSNI period of collection and female CVD mortality")+
ggrepel::geom_text_repel(data = outlier14, aes(label = location_name                                        ))  
  


```

##Univariable regression

###CVD mort (female) ~ GSNI

```{r}
m1 <- lm(cvd14fem ~ twobias, data = analysis14)
summary(m1)
plot(m1)

m1tidy <- tidy(m1)
kable(m1tidy, digits = 3)
```


###CVD mort (female) ~ physicians per 1000
```{r}
m2 <- lm(cvd14fem ~ phy14edit, data = analysis14)
summary(m2)
plot(m2)
```

###CVD mort (female) ~ mean years of schooling
```{r}
m3 <- lm(cvd14fem ~ scl14, data = analysis14)
summary(m3)
plot(m3)
```

###CVD mort (female) ~ GDP per capita
```{r}
m4 <- lm(cvd14fem ~ GDP14, data = analysis14)
summary(m4)
plot(m4)
```

###CVD mort (female) ~ MMR
```{r}
m5 <- lm(cvd14fem ~ MMR14, data = analysis14)
summary(m5)
plot(m5)
```

###CVD mort (female) ~ year of index
```{r}
m6 <- lm(cvd14fem ~ GSNI_PERIOD, data = analysis14)
summary(m6)
plot(m6)
```

##Multivariable regression
```{r}

m7 <- lm(cvd14fem ~ twobias + phy14edit + scl14+ MMR14+ GDP14, data = analysis14)
summary(m7)
plot(m7)

#added in GSNI PERIOD
m8 <- lm(cvd14fem ~ twobias + phy14edit + scl14+ MMR14+ GDP14+GSNI_PERIOD, data = analysis14)
summary(m8)
plot(m8)


m8tidy <- tidy(m8)
kable(m8tidy, digits = 3)
#r2 higher in m8

#without MMR
m7.1 <- lm(cvd14fem ~ twobias + phy14edit + scl14+ GDP14+GSNI_PERIOD, data = analysis14)
summary(m7.1)
plot(m7.1)

```

http://r-statistics.co/Outlier-Treatment-With-R.html
```{r}

cooksd <- cooks.distance(m8)

plot(cooksd, pch="*", cex=2, main="Influential Obs by Cooks distance")  # plot cook's distance
abline(h = 4*mean(cooksd, na.rm=T), col="red")  # add cutoff line
text(x=1:length(cooksd)+1, y=cooksd, labels=ifelse(cooksd>4*mean(cooksd, na.rm=T),names(cooksd),""), col="red")  # add labels

influential <- as.numeric(names(cooksd)[(cooksd > 4*mean(cooksd, na.rm=T))])
# influential row numbers
analysis14[influential, ]  # influential observations.

car::outlierTest(m8)
```


assess for colinearity
```{r}

car::vif(m7.1)
#all <4
```

assess for normality of residuals 
```{r}
shapiro.test(m7$residuals)


shapiro.test(m8$residuals)

#evidence of non-normality in the residuals
```
```{r}
#check for independence
plot(m7$residuals, type = "o")
plot(m8$residuals, type = "o")
```


####Assumptions
Linearity - residual vs fitted appears within limits
Normality of residuals - QQplots show some deviance, shapiro-wilk are significant suggesting lack of normality 
Independence - looks fine
multicolinearity - all within limits
homoscedacity - scale-location plot for m8 looks better than m7, (flatter and closer to 1)

####Try removing Uzbekistan
```{r}
analysis14noUzbek <- analysis14[-63,]

m8noUzbek <- lm(cvd14fem ~ twobias + phy14edit + scl14+ MMR14+ GDP14+GSNI_PERIOD, data = analysis14noUzbek)
summary(m8noUzbek)
plot(m8noUzbek)
shapiro.test(m8noUzbek$residuals)

#decreases coefficient slightly 4.8 to 4.2, now 46 (Quatar) is outside cook's distance... probably not better to exlude it
```



####Comparing AIC (using corrected as number of covariates is 10% of data https://stats.stackexchange.com/questions/86768/aicc-for-small-sample-sizes)

```{r}
AICcmodavg::AICc(m7)

AICcmodavg::AICc(m8) #m8 has lower AICc -> preferable
```


#OUTCOME 2 - CVD mortality male

##Scatter plots

###CVD mort (male) ~ GSNI

```{r}
ggplot(analysis14, aes(x = twobias, y = cvd14male)) +
  geom_point(color = "darkgreen")+
  geom_smooth(method = "lm")+
         labs(x = "percentage with 2 GSNI bias (%)",
                y = "male age standardised CVD mortaltiy rate (per 100k)",
                title = "Scatter plot of GSNI (2 bias) and male CVD mortality")

analysis14[,cor.test(twobias,cvd14male,method = "spearman")]

```


###CVD mort (male) ~ physicians per 1000

```{r}
ggplot(analysis14, aes(x = phy14edit, y = cvd14male)) +
  geom_point(color = "darkgreen")+
  geom_smooth(method = "lm")+
         labs(x = "physicians per 1000 population",
                y = "male age standardised CVD mortaltiy rate (per 100k)",
                title = "Scatter plot of phycisians per 1000 and male CVD mortality")

analysis14[,cor.test(cvd14male,phy14edit,method = "spearman")]

```

###CVD mort (male) ~ mean years of schooling

```{r}
ggplot(analysis14, aes(x = scl14, y = cvd14male)) +
  geom_point(color = "darkgreen")+
  geom_smooth(method = "lm")+
         labs(x = "mean years of schooling",
                y = "male age standardised CVD mortaltiy rate (per 100k)",
                title = "Mean years of schooling and male CVD mortality")

analysis14[,cor.test(scl14,cvd14male,method = "spearman")]

```

###CVD mort (male) ~ GDP per capita

```{r}
ggplot(analysis14, aes(x =GDP14, y = cvd14male)) +
  geom_point(color = "darkgreen")+
  geom_smooth(method = "lm")+
         labs(x = "GDP per capita",
                y = "male age standardised CVD mortaltiy rate (per 100k)",
                title = "GDP per capita and male CVD mortality")

analysis14[,cor.test(GDP14,cvd14male,method = "spearman")]
```



###CVD mort (male) ~ year of index
```{r}

outlier14male <- analysis14[cvd14male>750,] #set up to label extreme values
  
#boxplot separated by period of collection
ggplot(analysis14, aes(x = GSNI_PERIOD, y = cvd14male)) +
  geom_boxplot(color = "violetred3", notch= FALSE) +
         labs(x = "GSNI period",
                y = "male age standardised CVD mortaltiy rate",
                title = "GSNI period of collection and male CVD mortality")+
ggrepel::geom_text_repel(data = outlier14male, aes(label = location_name                                        ))  
  


```

##Univariable regression

###CVD mort (male) ~ GSNI

```{r}
m1m <- lm(cvd14male ~ twobias, data = analysis14)
summary(m1m)
plot(m1m)

m1mtidy <- tidy(m1m)
kable(m1mtidy, digits = 3)
```


###CVD mort (male) ~ physicians per 1000
```{r}
m2m <- lm(cvd14male ~ phy14edit, data = analysis14)
summary(m2m)
plot(m2m)
```

###CVD mort (male) ~ mean years of schooling
```{r}
m3m <- lm(cvd14male ~ scl14, data = analysis14)
summary(m3m)
plot(m3m)
```

###CVD mort (male) ~ GDP per capita
```{r}
m4m <- lm(cvd14male ~ GDP14, data = analysis14)
summary(m4m)
plot(m4m)
```


###CVD mort (male) ~ year of index
```{r}
m6m <- lm(cvd14male ~ GSNI_PERIOD, data = analysis14)
summary(m6m)
plot(m6m)
```

##Multivariable regression
```{r}

m7m <- lm(cvd14male ~ twobias + phy14edit + scl14 + GDP14 + GSNI_PERIOD, data = analysis14)
summary(m7m)
plot(m7m)

```

http://r-statistics.co/Outlier-Treatment-With-R.html



assess for colinearity
```{r}
car::vif(m7m)
#all <4
```



```{r}
#check for independence
plot(m7m$residuals, type = "o")

```


####Assumptions
Linearity - residual vs fitted appears within limits
Normality of residuals - QQplots appear within limits
Independence - within limits
multicolinearity - all within limits
homoscedacity - scale-location plot appears within limits



#OUTCOME 3 - Population CVD mortality

##Scatter plots

###CVD mort ~ GSNI
```{r}
ggplot(analysis14, aes(x = twobias, y = cvd14both)) +
  geom_point(color = "violetred3")+
  geom_smooth(method = "lm")+
         labs(x = "percentage with 2 GSNI bias (%)",
                y = "age standardised CVD mortaltiy rate (per 100k)",
                title = "Scatter plot of GSNI (2 bias) and CVD mortality (2014)")

analysis14[,cor.test(twobias,cvd14both,method = "spearman")]

```

###CVD mort  ~ physicians per 1000

```{r}
ggplot(analysis14, aes(x = phy14edit, y = cvd14both)) +
  geom_point(color = "violetred3")+
  geom_smooth(method = "lm")+
         labs(x = "physicians per 1000 population",
                y = "age standardised CVD mortaltiy rate (per 100k)",
                title = "Scatter plot of phycisians per 1000 and CVD mortality")

analysis14[,cor.test(cvd14both,phy14edit,method = "spearman")]

```
###CVD mort ~ mean years of schooling
```{r}
ggplot(analysis14, aes(x = scl14, y = cvd14both)) +
  geom_point(color = "violetred3")+
  geom_smooth(method = "lm")+
         labs(x = "mean years of schooling",
                y = "age standardised CVD mortaltiy rate (per 100k)",
                title = "Mean years of schooling and CVD mortality")

analysis14[,cor.test(scl14,cvd14both,method = "spearman")]

```
###CVD mort  ~ GDP per capita
```{r}
ggplot(analysis14, aes(x =GDP14, y = cvd14both)) +
  geom_point(color = "violetred3")+
  geom_smooth(method = "lm")+
         labs(x = "GDP per capita",
                y = "age standardised CVD mortaltiy rate (per 100k)",
                title = "GDP per capita and CVD mortality")

analysis14[,cor.test(GDP14,cvd14both,method = "spearman")]
```
###CVD mort  ~ MMR
```{r}
ggplot(analysis14, aes(x = MMR14, y = cvd14both)) +
  geom_point(color = "violetred3")+
  geom_smooth(method = "lm")+
         labs(x = "Maternal Mortality Ratio",
                y = "age standardised CVD mortaltiy rate (per 100k)",
                title = "MMR and CVD mortality")

analysis14[,cor.test(MMR14,cvd14both,method = "spearman")]

```
###CVD mort  ~ year of index
```{r}

outlier14.2 <- analysis14[cvd14both>600, ]
ggplot(analysis14, aes(x = GSNI_PERIOD, y = cvd14both)) +
  geom_boxplot(color = "violetred3", notch= TRUE)+
  
         labs(x = "GSNI period",
                y = "age standardised CVD mortaltiy rate (per 100k) (2014)",
                title = "GSNI period of collection and CVD mortality")+
 ggrepel::geom_text_repel(data = outlier14.2, aes(label = location_name                                        ))   


```

##Univariable regression

###CVD mort  ~ GSNI
```{r}
m9 <- lm(cvd14both ~ twobias, data = analysis14)
summary(m9)
plot(m9)

m9tidy <- tidy(m9)
kable(m9tidy, digits = 3)
```
###CVD mort  ~ physicians per 1000
```{r}
m10 <- lm(cvd14both ~ phy14edit, data = analysis14)
summary(m10)
plot(m10)
```
###CVD mort  ~ mean years of schooling
```{r}
m11 <- lm(cvd14both ~ scl14, data = analysis14)
summary(m11)
plot(m11)
```
###CVD mort  ~ GDP per capita
```{r}
m12 <- lm(cvd14both ~ GDP14, data = analysis14)
summary(m12)
plot(m12)
```
###CVD mort ~ MMR
```{r}
m13 <- lm(cvd14both ~ MMR14, data = analysis14)
summary(m13)
plot(m13)
```
###CVD mort ~ year of index
```{r}
m14 <- lm(cvd14both ~ GSNI_PERIOD, data = analysis14)
summary(m14)
plot(m14)
```
##Multivariable regression

```{r}

m15<- lm(cvd14both ~ twobias + phy14edit + scl14+ MMR14+ GDP14, data = analysis14)
summary(m15)
plot(m15)

#added in GSNI PERIOD
m16 <- lm(cvd14both ~ twobias + phy14edit + scl14+ MMR14+ GDP14+GSNI_PERIOD, data = analysis14)
summary(m16)
plot(m16)



car::outlierTest(m16)

#remove MMR
m16.1 <- lm(cvd14both ~ twobias + phy14edit + scl14+ GDP14+GSNI_PERIOD, data = analysis14)
summary(m16.1)
plot(m16.1)

```


assess for colinearity
```{r}

car::vif(m16.1)
#all <4
```







Assumptions
Linearity - residual vs fitted appears within limits
Normality of residuals - QQplots show some deviance, shapiro-wilk are significant suggesting lack of normality 
Independence - looks fine
multicolinearity - all within limits
homoscedacity - scale-location plot for m8 looks better than m7, (flatter and closer to 1)


#OUTCOME 4 - CVD mortality female/male ratio

##Scatter plots

###CVD ratio ~ GSNI

```{r}
ggplot(analysis14, aes(x = twobias, y = cvdratio14)) +
  geom_point(color = "orange")+
  geom_smooth(method = "lm")+
         labs(x = "percentage with 2 GSNI bias (%)",
                y = "female/male CVD mortaltiy ratio",
                title = "Scatter plot of GSNI (2 bias) and female/male CVD mortality ratio")

analysis14[,cor.test(twobias,cvdratio14,method = "spearman")]

```


###CVD ratio ~ physicians per 1000

```{r}
ggplot(analysis14, aes(x = phy14edit, y = cvdratio14)) +
  geom_point(color = "orange")+
  geom_smooth(method = "lm")+
         labs(x = "physicians per 1000 population",
                y = "female/male CVD mortaltiy ratio",
                title = "Scatter plot of phycisians per 1000 and CVD mortality ratio")

analysis14[,cor.test(cvdratio14,phy14edit,method = "spearman")]

```

###CVD ratio ~ mean years of schooling

```{r}
ggplot(analysis14, aes(x = scl14, y = cvdratio14)) +
  geom_point(color = "orange")+
  geom_smooth(method = "lm")+
         labs(x = "mean years of schooling",
                y = "femlae/male CVD mortaltiy ratio",
                title = "Mean years of schooling and CVD mortality ratio")

analysis14[,cor.test(scl14,cvdratio14,method = "spearman")]

```

###CVD ratio ~ GDP per capita

```{r}
ggplot(analysis14, aes(x =GDP14, y = cvdratio14)) +
  geom_point(color = "orange")+
  geom_smooth(method = "lm")+
         labs(x = "GDP per capita",
                y = "female/male CVD mortaltiy ratio",
                title = "GDP per capita and CVD mortality ratio")

analysis14[,cor.test(GDP14,cvdratio14,method = "spearman")]
```




##Univariable regression

###CVD ratio ~ GSNI

```{r}
m1r <- lm(cvdratio14 ~ twobias, data = analysis14)
summary(m1r)
plot(m1r)

```


###CVD ratio ~ physicians per 1000
```{r}
m2r <- lm(cvdratio14 ~ phy14edit, data = analysis14)
summary(m2r)
plot(m2r)
```

###CVD ratio ~ mean years of schooling
```{r}
m3r <- lm(cvdratio14 ~ scl14, data = analysis14)
summary(m3r)
plot(m3r)
```

###CVD ratio ~ GDP per capita
```{r}
m4r <- lm(cvdratio14 ~ GDP14, data = analysis14)
summary(m4r)
plot(m4r)
```


###CVD ratio ~ year of index
```{r}
m6r <- lm(cvdratio14 ~ GSNI_PERIOD, data = analysis14)
summary(m6r)
plot(m6r)
```

##Multivariable regression
```{r}

m7r <- lm(cvdratio14 ~ twobias + phy14edit + scl14 + GDP14 + GSNI_PERIOD, data = analysis14)
summary(m7r)
plot(m7r)

```

http://r-statistics.co/Outlier-Treatment-With-R.html

```{r}

cooksd <- cooks.distance(m7r)

plot(cooksd, pch="*", cex=2, main="Influential Obs by Cooks distance")  # plot cook's distance
abline(h = 4*mean(cooksd, na.rm=T), col="red")  # add cutoff line
text(x=1:length(cooksd)+1, y=cooksd, labels=ifelse(cooksd>4*mean(cooksd, na.rm=T),names(cooksd),""), col="red")  # add labels

influential <- as.numeric(names(cooksd)[(cooksd > 4*mean(cooksd, na.rm=T))])
# influential row numbers
analysis14[influential, ]  # influential observations.

car::outlierTest(m7r)

#Quator is outside Cooks distance

noQuatar <- analysis14[-52,]
m7rnoQ <- lm(cvdratio14 ~ twobias + phy14edit + scl14 + GDP14 + GSNI_PERIOD, data = noQuatar)
summary(m7rnoQ)
plot(m7rnoQ)

#after Quatar removed not significant

```

assess for colinearity
```{r}
car::vif(m7rnoQ)
#all <4
```



```{r}
#check for independence
plot(m7rnoQ$residuals, type = "o")

```


####Assumptions
Linearity - residual vs fitted appears within limits
Normality of residuals - QQplots appear within limits
Independence - within limits
multicolinearity - all within limits
homoscedacity - scale-location plot appears within limits

#OUTCOME 5 Life expectancy for women

## Scatter plots

### Life Expectancy ~ GSNI
```{r}
ggplot(analysis14, aes(x = twobias, y = LEbirth2015)) +
  geom_point(color = "violetred3")+
  geom_smooth(method = "lm")+
         labs(x = "percentage with 2 GSNI bias (%)",
                y = "Female life expectancy",
                title = "Scatter plot of GSNI (2 bias) and female life expectancy")

analysis14[,cor.test(twobias,LEbirth2015,method = "spearman")]

```

###Life Expectancy ~ physicians per 1000
```{r}
ggplot(analysis14, aes(x = phy14edit, y = LEbirth2015)) +
  geom_point(color = "violetred3")+
         labs(x = "physicians per 1000",
                y = "Female life expectancy",
                title = "Scatter plot of physicians per 1000 and female life expectancy")

analysis14[,cor.test(phy14edit,LEbirth2015,method = "spearman")]

```
###Life Expectancy ~ mean years of schooling
```{r}
ggplot(analysis14, aes(x = scl14, y = LEbirth2015)) +
  geom_point(color = "violetred3")+
  geom_smooth(method = "lm")+
         labs(x = "mean years of schooling",
                y = "Female life expectancy",
                title = "Scatter plot of mean years of schooling and female life expectancy")

analysis14[,cor.test(scl14,LEbirth2015,method = "spearman")]

```
###Life Expectancy ~ GDP per capita
```{r}
ggplot(analysis14, aes(x = GDP14, y = LEbirth2015)) +
  geom_point(color = "violetred3")+
         labs(x = "GDP per capita",
                y = "Female life expectancy",
                title = "GDP per capita and female life expectancy")

#can take a log of GDP has better relationship
ggplot(analysis14, aes(x = log(GDP14), y = LEbirth2015)) +
  geom_point(color = "violetred3")+
  geom_smooth(method = "lm")+
         labs(x = "log GDP per capita",
                y = "Female life expectancy",
                title = "log GDP per capita and female life expectancy")

analysis14[,cor.test(GDP14,LEbirth2015,method = "spearman")]

```
###Life Expectancy ~ MMR
```{r}
ggplot(analysis14, aes(x = MMR14, y = LEbirth2015)) +
  geom_point(color = "violetred3")+
  geom_smooth(method = "lm")+
         labs(x = "MMR",
                y = "Female life expectancy",
                title = "Scatter plot of MMR and female life expectancy")

analysis14[,cor.test(MMR14,LEbirth2015,method = "spearman")]

```
###Life Expectancy ~ year of index
```{r}
outlier14.3 <- analysis14[LEbirth2015<65,]

ggplot(analysis14, aes(x = GSNI_PERIOD, y = LEbirth2015)) +
  geom_boxplot(color = "violetred3", notch= TRUE)+
  
         labs(x = "GSNI period",
                y = "Life expectancy at birth (2015)",
                title = "GSNI period of collection and life expectancy")+
ggrepel::geom_text_repel(data = outlier14.3, aes(label = location_name                                        ))  

```


##Univariable regression

###Life Expectancy ~ GSNI
```{r}
m17 <- lm(LEbirth2015 ~ twobias, data = analysis14)
summary(m17)
plot(m17)

m17tidy <- tidy(m17)
kable(m17tidy, digits = 3)
```

###Life Expectancy ~ physicians per 1000
```{r}
m18 <- lm(LEbirth2015 ~ phy14edit, data = analysis14)
summary(m18)
plot(m18)
```

###Life Expectancy ~ mean years of schooling
```{r}
m19 <- lm(LEbirth2015 ~ scl14, data = analysis14)
summary(m19)
plot(m19)
```

###Life Expectancy ~ GDP per capita
```{r}
m20 <- lm(LEbirth2015 ~ GDP14, data = analysis14)
summary(m20)
plot(m20)
```
###Life Expectancy ~ MMR
```{r}
m21 <- lm(LEbirth2015 ~ MMR14, data = analysis14)
summary(m21)
plot(m21)
```
###Life Expectancy ~ year of index
```{r}
m22 <- lm(LEbirth2015 ~ GSNI_PERIOD, data = analysis14)
summary(m22)
plot(m22)
```

##Multivariable regression
```{r}

m23 <- lm(LEbirth2015 ~ twobias + phy14edit + scl14+ MMR14+ GDP14, data = analysis14)
summary(m23)
plot(m23)

#added in GSNI PERIOD
m24 <- lm(LEbirth2015 ~ twobias + phy14edit + scl14+ MMR14+ GDP14+GSNI_PERIOD, data = analysis14)
summary(m24)
plot(m24)

m24tidy <- tidy(m24)
kable(m24tidy, digits = 3)

car::outlierTest(m24)

m24.2 <- lm(LEbirth2015 ~ twobias + phy14edit + scl14+ GDP14+GSNI_PERIOD, data = analysis14)
summary(m24.2)
plot(m24.2)

#GDP is not linear with LE, https://kenbenoit.net/assets/courses/ME104/logmodels2.pdf
#therefore can try a log also with phy based on plotting hte regression terms of the model below

m24.3 <- lm(LEbirth2015 ~ twobias + log(phy14edit) + scl14+ log(GDP14)+GSNI_PERIOD +MMR14, data = analysis14)
summary(m24.3)
plot(m24.3)

```


```{r}

#plot regression terms of the model
par(mfrow = c(2, 3)) # We have 4 predictors
termplot(m24.3, partial.resid = TRUE)
```



assess for colinearity
```{r}
car::vif(m24.3)
#log(phy14edit is 4.8)
```

assess for normality of residuals 





Assumptions
Linearity - residual vs fitted appears non linear, will investigate log transformation of variables
Normality of residuals - QQplots look ok and shapiro wilk not significant
Independence - looks fine
multicolinearity - all <5
homoscedacity - scale-location plot looks ok around 1

####no Nigeria
```{r}
analysis14noNig <- analysis14[-40,]
m24noNig <- lm(LEbirth2015 ~ twobias + log(phy14edit) + scl14+ MMR14+ log(GDP14) + GSNI_PERIOD, data = analysis14noNig)
summary(m24noNig)
plot(m24noNig)

shapiro.test(m24noNig$residuals)

#removing Nigeria doesn't change results
```



Comparing AIC (using corrected as number of covariates is 10% of data https://stats.stackexchange.com/questions/86768/aicc-for-small-sample-sizes)

```{r}
AICcmodavg::AICc(m23)

AICcmodavg::AICc(m24) #m24 has lower AICc -> preferable

AICcmodavg::AICc(m24noNig)

```

#OUTCOME 6 Life expectancy for men

## Scatter plots

### Life Expectancy ~ GSNI
```{r}
ggplot(analysis14, aes(x = twobias, y = MLEbirth2015)) +
  geom_point(color = "violetred3")+
  geom_smooth(method = "lm")+
         labs(x = "percentage with 2 GSNI bias (%)",
                y = "male life expectancy",
                title = "Scatter plot of GSNI (2 bias) and male life expectancy")

analysis14[,cor.test(twobias,MLEbirth2015,method = "spearman")]

```

###Life Expectancy ~ physicians per 1000
```{r}
ggplot(analysis14, aes(x = phy14edit, y = MLEbirth2015)) +
  geom_point(color = "violetred3")+
         labs(x = "physicians per 1000",
                y = "male life expectancy",
                title = "Scatter plot of physicians per 1000 and male life expectancy")

analysis14[,cor.test(phy14edit,MLEbirth2015,method = "spearman")]

```
###Life Expectancy ~ mean years of schooling
```{r}
ggplot(analysis14, aes(x = scl14, y = MLEbirth2015)) +
  geom_point(color = "violetred3")+
  geom_smooth(method = "lm")+
         labs(x = "mean years of schooling",
                y = "male life expectancy",
                title = "Scatter plot of mean years of schooling and male life expectancy")

analysis14[,cor.test(scl14,MLEbirth2015,method = "spearman")]

```
###Life Expectancy ~ GDP per capita
```{r}
ggplot(analysis14, aes(x = GDP14, y = MLEbirth2015)) +
  geom_point(color = "violetred3")+
         labs(x = "GDP per capita",
                y = "male life expectancy",
                title = "GDP per capita and male life expectancy")

#can take a log of GDP has better relationship
ggplot(analysis14, aes(x = log(GDP14), y = MLEbirth2015)) +
  geom_point(color = "violetred3")+
  geom_smooth(method = "lm")+
         labs(x = "log GDP per capita",
                y = "male life expectancy",
                title = "log GDP per capita and male life expectancy")

analysis14[,cor.test(GDP14,MLEbirth2015,method = "spearman")]

```

###Life Expectancy ~ year of index
```{r}
outlier14.3m <- analysis14[MLEbirth2015<65,]

ggplot(analysis14, aes(x = GSNI_PERIOD, y = MLEbirth2015)) +
  geom_boxplot(color = "violetred3", notch= TRUE)+
  
         labs(x = "GSNI period",
                y = "Male life expectancy at birth (2015)",
                title = "GSNI period of collection and life expectancy")+
ggrepel::geom_text_repel(data = outlier14.3m, aes(label = location_name                                        ))  

```


##Univariable regression

###Life Expectancy ~ GSNI
```{r}
m17m <- lm(MLEbirth2015 ~ twobias, data = analysis14)
summary(m17m)
plot(m17m)

```

###Life Expectancy ~ physicians per 1000
```{r}
m18m <- lm(MLEbirth2015 ~ phy14edit, data = analysis14)
summary(m18m)
plot(m18m)
```

###Life Expectancy ~ mean years of schooling
```{r}
m19m <- lm(MLEbirth2015 ~ scl14, data = analysis14)
summary(m19m)
plot(m19m)
```

###Life Expectancy ~ GDP per capita
```{r}
m20m <- lm(MLEbirth2015 ~ GDP14, data = analysis14)
summary(m20m)
plot(m20m)
```

###Life Expectancy ~ year of index
```{r}
m22m <- lm(MLEbirth2015 ~ GSNI_PERIOD, data = analysis14)
summary(m22m)
plot(m22m)
```

##Multivariable regression
```{r}


m24m <- lm(MLEbirth2015 ~ twobias + log(phy14edit) + scl14 + log(GDP14) + GSNI_PERIOD, data = analysis14)
summary(m24m)
plot(m24m)

```






assess for colinearity
```{r}

car::vif(m24m)
#all <4
```

assess for normality of residuals 



```{r}
#check for independence

plot(m24m$residuals, type = "o")
```


Assumptions
Linearity - residual vs fitted appears non linear, will investigate log transformation of variables
Normality of residuals - QQplots look ok and shapiro wilk not significant
Independence - looks fine
multicolinearity - all within limits
homoscedacity - scale-location plot looks ok around 1




#---

