---
title: "dissertation analysis"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(qwraps2)
options(qwraps2_markup = "markdown")
library(readr)
library(readxl)
library(data.table)
library(ggplot2)
library(dplyr)
library(skimr)
library(tidyverse)
library(broom)
library(knitr)
library(SmartEDA)

analysis17 <- read_csv("analysisratio17.csv")
setDT(analysis17)

analysis17[, GSNI_PERIOD := as.factor(GSNI_PERIOD)]
analysis17[,"...1":= NULL]
```


#2017 

# CODE BOOK (analysis17)

+-------------------------+----------------------------------------------------+
| Variable                | Description                                        |
+=========================+====================================================+
| location_name           | country                                            |
+-------------------------+----------------------------------------------------+
| GSNI_PERIOD             | year data collection took place for GSNI survey    |
|                         |  2005-2009 or 2010-2014                            |
+-------------------------+----------------------------------------------------+
| twobias                 | share of people with at least two bias             |
+-------------------------+----------------------------------------------------+
| cvd17fem                | age standardised CVD mortality rates for women per |            |                         |    100,000 population                              |
|                         |    (from global burden of disease)                 |            
|                         |     2017                                           |
+-------------------------+----------------------------------------------------+
| cvd17male               | as above for whole males                           |
+-------------------------+----------------------------------------------------+
| cvd17both               | as above for whole population                      |
+-------------------------+----------------------------------------------------+
| cvdratio17              | ratio female/male CVD mortality rates              |
+-------------------------+----------------------------------------------------+
| LEbirth2019             |  life expectancy at birth for                      |         
|                         |  females in 2015, from WHO (years)                 |
+-------------------------+----------------------------------------------------+
| MLEbirth2019            |  life expectancy at birth for                      |         
|                         |  males in 2015, from WHO (years)                   |
+-------------------------+----------------------------------------------------+
| phy17edit               | physicians per 1000 population 2014 (world bank)   |
|                         |  data added from other years n = 19 (2013-17)      |
+-------------------------+----------------------------------------------------+
| scl17                   | mean years of school 2017 (ourworldindata)         |
+-------------------------+----------------------------------------------------+
| GDP17                   | Gross domestic product per capita (current US      |            |                         | dollar)  2017  (world bank)                        |
+-------------------------+----------------------------------------------------+
| MMR17                   | Maternal Mortality Ratio 2017   (world bank)       |
+-------------------------+----------------------------------------------------+

#Descriptive analysis

```{r}
glimpse(analysis17)
summary(analysis17)
```

```{r}

analysis17copy <- copy(analysis17)
analysis17copy2 <- copy(analysis17)

#create data table without MMR
analysis17noMMR <- cbind(analysis17copy[, MMR17:=NULL])

#create table without MMR and without NAs
analysis17noMMRnoNA <- cbind(na.exclude(analysis17noMMR))

#creat table with MMR with no NAs
analysis17noNA <- cbind(na.exclude(analysis17copy2)) #exclude rows with NA as will prevent model comparison

#just use analysis14 for all analysis as will exclude NAs

```

list of countries included
```{r}
analysis17noNA[,location_name] #66 rows (66 data points for MMR analysis)
analysis17noMMRnoNA[,location_name] #67 rows 
analysis17[,location_name] #75 rows
```




#####Histograms
```{r}
#ggplot(analysis17noNA, aes(x=twobias))+geom_histogram(binwidth=5, color = "Black", fill = "Pink")

analysis17[,hist(twobias)]
analysis17[,hist(cvd17fem)]
analysis17[,hist(cvd17male)]
analysis17[,hist(cvdratio17)]
analysis17[,hist(cvd17both)]
analysis17[,hist(LEbirth2019)]
analysis17[,hist(MLEbirth2019)]
analysis17[,hist(phy17edit)]
analysis17[,hist(scl17)]
analysis17[,hist(GDP17)]
analysis17[,hist(MMR17)]

#no continuous variables variables normally distributed

summary(analysis17 
        )
```

#####summary tables
https://statsandr.com/blog/descriptive-statistics-in-r/
https://cran.r-project.org/web/packages/qwraps2/vignettes/summary-statistics.html

```{r}
our_summary2 <-
  list("GSNI index 2 or more bias" =
       list("min"       = ~ round(min(twobias),1),
            "max"       = ~ round(max(twobias),1),
            "median (IQR)" = ~ median_iqr(twobias)),
       "Female CVD age adjusted mortality rate (per 100 000 population)" =
       list("min"       = ~ round(min(cvd17fem),1),
            "max"       = ~ round(max(cvd17fem),1),
            "median (IQR)" = ~ median_iqr(cvd17fem)),
       "Population CVD age adjusted mortality rate (per 100 000 population)" =
       list("min"       = ~ round(min(cvd17both),1),
            "max"       = ~ round(max(cvd17both),1),
            "median (IQR)" = ~ median_iqr(cvd17both)),
       "Female life expectancy at birth" =
       list("min"       = ~ round(min(LEbirth2019),1),
            "max"       = ~ round(max(LEbirth2019),1),
            "median (IQR)" = ~ median_iqr(LEbirth2019)),
       "Physicians per 1000 population" =
       list("min"       = ~ round(min(phy17edit),1),
            "max"       = ~ round(max(phy17edit),1),
            "median (IQR)" = ~ median_iqr(phy17edit)),
       "GDP per capita" =
       list("min"       = ~ round(min(GDP17),1),
            "max"       = ~ round(max(GDP17),1),
            "median (IQR)" = ~ median_iqr(GDP17)),
        "Maternal Mortality Ratio" =
       list("min"       = ~ round(min(MMR17),1),
            "max"       = ~ round(max(MMR17),1),
            "median (IQR)" = ~ median_iqr(MMR17)),
       "Mean years of schooling" =
       list("min"       = ~ round(min(scl17),1),
            "max"       = ~ round(max(scl17),1),
            "median (IQR)" = ~ median_iqr(scl17)),
       "Year of Index Collection" =
       list("2005-2009" = ~ n_perc0(GSNI_PERIOD == levels(analysis17$GSNI_PERIOD)[1]),
            "2010-2017"  = ~ n_perc0(GSNI_PERIOD == levels(analysis17$GSNI_PERIOD)[2]))
)
            
```





```{r}

whole17 <- summary_table(analysis17, our_summary2)
whole17
```

```{r}
our_summary2 <-
  list("GSNI index 2 or more bias" =
       list("min"       = ~ round(min(twobias),1),
            "max"       = ~ round(max(twobias),1),
            "median (IQR)" = ~ median_iqr(twobias),
        "missing" = ~sum(is.na(twobias))),
       "Female CVD age adjusted mortality rate (per 100 000 population)" =
       list("min"       = ~ round(min(cvd17fem),1),
            "max"       = ~ round(max(cvd17fem),1),
            "median (IQR)" = ~ median_iqr(cvd17fem),
             "missing" = ~sum(is.na(cvd17fem))),
       "Male CVD age adjusted mortality rate (per 100 000 population)" =
       list("min"       = ~ round(min(cvd17male),1),
            "max"       = ~ round(max(cvd17male),1),
            "median (IQR)" = ~ median_iqr(cvd17male),
             "missing" = ~sum(is.na(cvd17male))),
       "Population CVD age adjusted mortality rate (per 100 000 population)" =
       list("min"       = ~ round(min(cvd17both),1),
            "max"       = ~ round(max(cvd17both),1),
            "median (IQR)" = ~ median_iqr(cvd17both),
             "missing" = ~sum(is.na(cvd17both))),
       "Ratio of female to male CVD mortality rates" =
       list("min"       = ~ round(min(cvdratio17),1),
            "max"       = ~ round(max(cvdratio17),1),
            "median (IQR)" = ~ median_iqr(cvdratio17),
             "missing" = ~sum(is.na(cvdratio17))),
       "Female life expectancy at birth" =
       list("min"       = ~ round(min(LEbirth2019),1),
            "max"       = ~ round(max(LEbirth2019),1),
            "median (IQR)" = ~ median_iqr(LEbirth2019),
             "missing" = ~sum(is.na(LEbirth2019))),
        "Male life expectancy at birth" =
       list("min"       = ~ round(min(MLEbirth2019),1),
            "max"       = ~ round(max(MLEbirth2019),1),
            "median (IQR)" = ~ median_iqr(MLEbirth2019),
             "missing" = ~sum(is.na(MLEbirth2019))),
       "Physicians per 1000 population" =
       list("min"       = ~ round(min(phy17edit, na.rm= TRUE),1), 
            "max"       = ~ round(max(phy17edit, na.rm= TRUE),1), 
            "median (IQR)" = ~ median_iqr(phy17edit, na_rm = TRUE),
             "missing" = ~sum(is.na(phy17edit))),
       "GDP per capita" =
       list("min"       = ~ round(min(GDP17, na.rm= TRUE),1),
            "max"       = ~ round(max(GDP17, na.rm= TRUE),1),
            "median (IQR)" = ~ median_iqr(GDP17, na_rm = TRUE),
             "missing" = ~sum(is.na(GDP17))),
        "Maternal Mortality Ratio" =
       list("min"       = ~ round(min(MMR17, na.rm= TRUE),1),
            "max"       = ~ round(max(MMR17, na.rm= TRUE),1),
            "median (IQR)" = ~ median_iqr(MMR17, na_rm = TRUE),
             "missing" = ~sum(is.na(MMR17))),
       "Mean years of schooling" =
       list("min"       = ~ round(min(scl17, na.rm= TRUE),1),
            "max"       = ~ round(max(scl17, na.rm= TRUE),1),
            "median (IQR)" = ~ median_iqr(scl17, na_rm = TRUE),
             "missing" = ~sum(is.na(scl17))),
       "Year of Index Collection" =
       list("2005-2009" = ~ n_perc0(GSNI_PERIOD == levels(analysis17$GSNI_PERIOD)[1]),
            "2010-2014"  = ~ n_perc0(GSNI_PERIOD == levels(analysis17$GSNI_PERIOD)[2])))

            
```






```{r}

whole17 <- summary_table(analysis17, our_summary2)
print(whole17,
      rtitle = "Summary Statistics",
      cnames = "n = 75")
```

#OUTCOME 1 - CVD mortality female

##Scatter plots

###CVD mort (female) ~ GSNI

```{r}
ggplot(analysis17, aes(x = twobias, y = cvd17fem)) +
  geom_point(color = "violetred3")+
  geom_smooth(method = "lm")+
         labs(x = "percentage with 2 GSNI bias (%)",
                y = "female age standardised CVD mortaltiy rate (per 100k)",
                title = "Scatter plot of GSNI (2 bias) and female CVD mortality")

analysis17[,cor.test(twobias,cvd17fem,method = "spearman")]

```


###CVD mort (female) ~ physicians per 1000

```{r}
ggplot(analysis17, aes(x = phy17edit, y = cvd17fem)) +
  geom_point(color = "violetred3")+
  geom_smooth(method = "lm")+
         labs(x = "physicians per 1000 population",
                y = "female age standardised CVD mortaltiy rate (per 100k)",
                title = "Scatter plot of phycisians per 1000 and female CVD mortality")

analysis17[,cor.test(cvd17fem,phy17edit,method = "spearman")]

```

###CVD mort (female) ~ mean years of schooling

```{r}
ggplot(analysis17, aes(x = scl17, y = cvd17fem)) +
  geom_point(color = "violetred3")+
  geom_smooth(method = "lm")+
         labs(x = "mean years of schooling",
                y = "female age standardised CVD mortaltiy rate (per 100k)",
                title = "Mean years of schooling and female CVD mortality")

analysis17[,cor.test(scl17,cvd17fem,method = "spearman")]

```

###CVD mort (female) ~ GDP per capita

```{r}
ggplot(analysis17, aes(x =GDP17, y = cvd17fem)) +
  geom_point(color = "violetred3")+
  geom_smooth(method = "lm")+
         labs(x = "GDP per capita",
                y = "female age standardised CVD mortaltiy rate (per 100k)",
                title = "GDP per capita and female CVD mortality")

analysis17[,cor.test(GDP17,cvd17fem,method = "spearman")]
```

###CVD mort (female) ~ MMR
```{r}
ggplot(analysis17, aes(x = MMR17, y = cvd17fem)) +
  geom_point(color = "violetred3")+
  geom_smooth(method = "lm")+
         labs(x = "Maternal Mortality Ratio",
                y = "female age standardised CVD mortaltiy rate (per 100k)",
                title = "MMR and female CVD mortality")

analysis17[,cor.test(MMR17,cvd17fem,method = "spearman")]

```

###CVD mort (female) ~ year of index
```{r}
outlier17.1<- analysis17[cvd17fem>550,]
ggplot(analysis17, aes(x = GSNI_PERIOD, y = cvd17fem)) +
  geom_boxplot(color = "violetred3", notch= TRUE)+
  
         labs(x = "GSNI period",
                y = "female age standardised CVD mortaltiy rate (per 100k) (2017)",
                title = "GSNI period of collection and female CVD mortality")+
ggrepel::geom_text_repel(data = outlier17.1, aes(label = location_name                                        ))  

```

##Univariable regression

###CVD mort (female) ~ GSNI

```{r}
m17.1 <- lm(cvd17fem ~ twobias, data = analysis17)
summary(m17.1)
plot(m17.1)
```


###CVD mort (female) ~ physicians per 1000
```{r}
m17.2 <- lm(cvd17fem ~ phy17edit, data = analysis17)
summary(m17.2)
plot(m17.2)
```

###CVD mort (female) ~ mean years of schooling
```{r}
m17.3 <- lm(cvd17fem ~ scl17, data = analysis17)
summary(m17.3)
plot(m17.3)
```

###CVD mort (female) ~ GDP per capita
```{r}
m17.4 <- lm(cvd17fem ~ GDP17, data = analysis17)
summary(m17.4)
plot(m17.4)
```

###CVD mort (female) ~ MMR
```{r}
m17.5 <- lm(cvd17fem ~ MMR17, data = analysis17)
summary(m17.5)
plot(m17.5)
```

###CVD mort (female) ~ year of index
```{r}
m17.6 <- lm(cvd17fem ~ GSNI_PERIOD, data = analysis17)
summary(m17.6)
plot(m17.6)
```

##Multivariable regression
```{r}

m17.7 <- lm(cvd17fem ~ twobias + phy17edit + scl17 + GDP17 + GSNI_PERIOD, data = analysis17)
summary(m17.7)
plot(m17.7)




```


assess for colinearity
```{r}
car::vif(m17.7)

#all <4
```


```{r}
#check for independence
plot(m17.7$residuals, type = "o")

```


#Assumptions
Linearity - residual vs fitted appears within limits
Normality of residuals - QQplots 
Independence - looks fine
multicolinearity - all within limits
homoscedacity - scale-location plots look ok

#OUTCOME 2 - male CVD mortality

##Scatter plots

###CVD mort ~ GSNI
```{r}
ggplot(analysis17, aes(x = twobias, y = cvd17male)) +
  geom_point(color = "violetred3")+
  geom_smooth(method = "lm")+
         labs(x = "percentage with 2 GSNI bias (%)",
                y = "age standardised male CVD mortaltiy rate (per 100k)",
                title = "Scatter plot of GSNI (2 bias) and maleCVD mortality")

analysis17[,cor.test(twobias,cvd17male,method = "spearman")]

```

###CVD mort  ~ physicians per 1000

```{r}
ggplot(analysis17, aes(x = phy17edit, y = cvd17male)) +
  geom_point(color = "violetred3")+
  geom_smooth(method = "lm")+
         labs(x = "physicians per 1000 population",
                y = "age standardised male CVD mortaltiy rate (per 100k)",
                title = "Scatter plot of phycisians per 1000 and male CVD mortality")

analysis17[,cor.test(cvd17male,phy17edit,method = "spearman")]

```
###CVD mort ~ mean years of schooling
```{r}
ggplot(analysis17, aes(x = scl17, y = cvd17male)) +
  geom_point(color = "violetred3")+
  geom_smooth(method = "lm")+
         labs(x = "mean years of schooling",
                y = "age standardised male CVD mortaltiy rate (per 100k)",
                title = "Mean years of schooling and CVD mortality")

analysis17[,cor.test(scl17,cvd17male,method = "spearman")]

```
###CVD mort  ~ GDP per capita
```{r}
ggplot(analysis17, aes(x =GDP17, y = cvd17male)) +
  geom_point(color = "violetred3")+
  geom_smooth(method = "lm")+
         labs(x = "GDP per capita",
                y = "age standardised male CVD mortaltiy rate (per 100k)",
                title = "GDP per capita and CVD mortality")

analysis17[,cor.test(GDP17,cvd17male,method = "spearman")]
```

###CVD mort  ~ year of index
```{r}
outlier17.2m <- analysis17[cvd17male>650,]
ggplot(analysis17, aes(x = GSNI_PERIOD, y = cvd17male)) +
  geom_boxplot(color = "violetred3", notch= TRUE)+
  
         labs(x = "GSNI period",
                y = "age standardised maleCVD mortaltiy rate (per 100k) (2017)",
                title = "GSNI period of collection and CVD mortality")+
ggrepel::geom_text_repel(data = outlier17.2m, aes(label = location_name                                        ))  
```

##Univariable regression

###CVD mort  ~ GSNI
```{r}
m17.9m <- lm(cvd17male ~ twobias, data = analysis17)
summary(m17.9m)
plot(m17.9m)
```
###CVD mort  ~ physicians per 1000
```{r}
m17.10m<- lm(cvd17male ~ phy17edit, data = analysis17)
summary(m17.10m)
plot(m17.10m)
```
###CVD mort  ~ mean years of schooling
```{r}
m17.11m <- lm(cvd17male ~ scl17, data = analysis17)
summary(m17.11m)
plot(m17.11m)
```
###CVD mort  ~ GDP per capita
```{r}
m17.1m2 <- lm(cvd17male ~ GDP17, data = analysis17)
summary(m17.12m)
plot(m17.12m)
```
###CVD mort ~ MMR
```{r}
m17.13m <- lm(cvd17male ~ MMR17, data = analysis17)
summary(m17.13m)
plot(m17.13m)
```
###CVD mort ~ year of index
```{r}
m17.14m <- lm(cvd17male ~ GSNI_PERIOD, data = analysis17)
summary(m17.14m)
plot(m17.14m)
```

##Multivariable regression

```{r}

m17.16m <- lm(cvd17male ~ twobias + phy17edit + scl17 + GDP17+GSNI_PERIOD, data = analysis17)
summary(m17.16m)
plot(m17.16m)


```


assess for colinearity
```{r}
car::vif(m17.16m)
#all <4
```


#Assumptions
Linearity - residual vs fitted appears within limits
Normality of residuals - QQplots show some deviance, shapiro-wilk are significant suggesting lack of normality 
Independence - looks fine
multicolinearity - all within limits
homoscedacity - scale-location plot looks ok



#OUTCOME 3 - population CVD mortality

##Scatter plots

###CVD mort ~ GSNI
```{r}
ggplot(analysis17, aes(x = twobias, y = cvd17both)) +
  geom_point(color = "violetred3")+
  geom_smooth(method = "lm")+
         labs(x = "percentage with 2 GSNI bias (%)",
                y = "age standardised CVD mortaltiy rate (per 100k)",
                title = "Scatter plot of GSNI (2 bias) and CVD mortality")

analysis17[,cor.test(twobias,cvd17both,method = "spearman")]

```

###CVD mort  ~ physicians per 1000

```{r}
ggplot(analysis17, aes(x = phy17edit, y = cvd17both)) +
  geom_point(color = "violetred3")+
  geom_smooth(method = "lm")+
         labs(x = "physicians per 1000 population",
                y = "age standardised CVD mortaltiy rate (per 100k)",
                title = "Scatter plot of phycisians per 1000 and CVD mortality")

analysis17[,cor.test(cvd17both,phy17edit,method = "spearman")]

```
###CVD mort ~ mean years of schooling
```{r}
ggplot(analysis17, aes(x = scl17, y = cvd17both)) +
  geom_point(color = "violetred3")+
  geom_smooth(method = "lm")+
         labs(x = "mean years of schooling",
                y = "age standardised CVD mortaltiy rate (per 100k)",
                title = "Mean years of schooling and CVD mortality")

analysis17[,cor.test(scl17,cvd17both,method = "spearman")]

```
###CVD mort  ~ GDP per capita
```{r}
ggplot(analysis17, aes(x =GDP17, y = cvd17both)) +
  geom_point(color = "violetred3")+
  geom_smooth(method = "lm")+
         labs(x = "GDP per capita",
                y = "age standardised CVD mortaltiy rate (per 100k)",
                title = "GDP per capita and CVD mortality")

analysis17[,cor.test(GDP17,cvd17both,method = "spearman")]
```
###CVD mort  ~ MMR
```{r}
ggplot(analysis17, aes(x = MMR17, y = cvd17both)) +
  geom_point(color = "violetred3")+
  geom_smooth(method = "lm")+
         labs(x = "Maternal Mortality Ratio",
                y = "age standardised CVD mortaltiy rate (per 100k)",
                title = "MMR and CVD mortality")

analysis17[,cor.test(MMR17,cvd17both,method = "spearman")]

```
###CVD mort  ~ year of index
```{r}
outlier17.2 <- analysis17[cvd17both>650,]
ggplot(analysis17, aes(x = GSNI_PERIOD, y = cvd17both)) +
  geom_boxplot(color = "violetred3", notch= TRUE)+
  
         labs(x = "GSNI period",
                y = "age standardised CVD mortaltiy rate (per 100k) (2017)",
                title = "GSNI period of collection and CVD mortality")+
ggrepel::geom_text_repel(data = outlier17.2, aes(label = location_name                                        ))  
```

##Univariable regression

###CVD mort  ~ GSNI
```{r}
m17.9 <- lm(cvd17both ~ twobias, data = analysis17)
summary(m17.9)
plot(m17.9)
```
###CVD mort  ~ physicians per 1000
```{r}
m17.10 <- lm(cvd17both ~ phy17edit, data = analysis17)
summary(m17.10)
plot(m17.10)
```
###CVD mort  ~ mean years of schooling
```{r}
m17.11 <- lm(cvd17both ~ scl17, data = analysis17)
summary(m17.11)
plot(m17.11)
```
###CVD mort  ~ GDP per capita
```{r}
m17.12 <- lm(cvd17both ~ GDP17, data = analysis17)
summary(m17.12)
plot(m17.12)
```
###CVD mort ~ MMR
```{r}
m17.13 <- lm(cvd17both ~ MMR17, data = analysis17)
summary(m17.13)
plot(m17.13)
```
###CVD mort ~ year of index
```{r}
m17.14 <- lm(cvd17both ~ GSNI_PERIOD, data = analysis17)
summary(m17.14)
plot(m17.14)
```

##Multivariable regression

```{r}

m17.16 <- lm(cvd17both ~ twobias + phy17edit + scl17 + GDP17+GSNI_PERIOD, data = analysis17)
summary(m17.16)
plot(m17.16)


```


assess for colinearity
```{r}
car::vif(m17.16)
#all <4
```


#Assumptions
Linearity - residual vs fitted appears within limits
Normality of residuals - QQplots show some deviance, shapiro-wilk are significant suggesting lack of normality 
Independence - looks fine
multicolinearity - all within limits
homoscedacity - scale-location plot looks ok


#OUTCOME 4 - ratio female/male CVD mortality

##Scatter plots

###CVD ratio ~ GSNI
```{r}
ggplot(analysis17, aes(x = twobias, y = cvdratio17)) +
  geom_point(color = "violetred3")+
  geom_smooth(method = "lm")+
         labs(x = "percentage with 2 GSNI bias (%)",
                y = "CVD mortaltiy ratio)",
                title = "Scatter plot of GSNI (2 bias) and CVD mortality rato")

analysis17[,cor.test(twobias,cvdratio17,method = "spearman")]

```

###CVD ratio  ~ physicians per 1000

```{r}
ggplot(analysis17, aes(x = phy17edit, y = cvdratio17)) +
  geom_point(color = "violetred3")+
  geom_smooth(method = "lm")+
         labs(x = "physicians per 1000 population",
                y = "CVD mortaltiy ratio",
                title = "Scatter plot of phycisians per 1000 and CVD ratio")

analysis17[,cor.test(cvdratio17,phy17edit,method = "spearman")]

```
###CVD ratio ~ mean years of schooling
```{r}
ggplot(analysis17, aes(x = scl17, y = cvdratio17)) +
  geom_point(color = "violetred3")+
  geom_smooth(method = "lm")+
         labs(x = "mean years of schooling",
                y = "CVD mortaltiy ratio",
                title = "Mean years of schooling and CVD ratio")

analysis17[,cor.test(scl17,cvdratio17,method = "spearman")]

```
###CVD ratui  ~ GDP per capita
```{r}
ggplot(analysis17, aes(x =GDP17, y = cvdratio17)) +
  geom_point(color = "violetred3")+
  geom_smooth(method = "lm")+
         labs(x = "GDP per capita",
                y = "CVD ratio",
                title = "GDP per capita and CVD ratio")

analysis17[,cor.test(GDP17,cvdratio17,method = "spearman")]
```

###CVD mort  ~ year of index
```{r}

ggplot(analysis17, aes(x = GSNI_PERIOD, y = cvdratio17)) +
  geom_boxplot(color = "violetred3", notch= TRUE)+
  
         labs(x = "GSNI period",
                y = "CVD mortaltiy ratio",
                title = "GSNI period of collection and CVD ratio"
)  
```

##Univariable regression

###CVD ratio  ~ GSNI
```{r}
m17.9r <- lm(cvdratio17 ~ twobias, data = analysis17)
summary(m17.9r)
plot(m17.9r)
```
###CVD ratio  ~ physicians per 1000
```{r}
m17.10r<- lm(cvdratio17 ~ phy17edit, data = analysis17)
summary(m17.10r)
plot(m17.10r)
```
###CVD ratio  ~ mean years of schooling
```{r}
m17.11r <- lm(cvdratio17 ~ scl17, data = analysis17)
summary(m17.11r)
plot(m17.11r)
```
###CVD ratio  ~ GDP per capita
```{r}
m17.12r <- lm(cvdratio17 ~ GDP17, data = analysis17)
summary(m17.12r)
plot(m17.12r)
```

###CVD ratio ~ year of index
```{r}
m17.14r <- lm(cvdratio17 ~ GSNI_PERIOD, data = analysis17)
summary(m17.14r)
plot(m17.14r)
```

##Multivariable regression

```{r}

m17.16r <- lm(cvdratio17 ~ twobias + phy17edit + scl17 + GDP17 + GSNI_PERIOD, data = analysis17)
summary(m17.16r)
plot(m17.16r)


```
```{r}


#Qatar is outside Cooks distance

noQatar17 <- analysis17[-52,]
m17.16rnoQ <- lm(cvdratio17 ~ twobias + phy17edit + scl17 + GDP17 + GSNI_PERIOD, data = noQatar17)
summary(m17.16rnoQ)
plot(m17.16rnoQ)

#after Quatar removed not significant

```

assess for colinearity
```{r}
car::vif(m17.16rnoQ)
#all <4
```


#Assumptions
Linearity - residual vs fitted appears within limits
Normality of residuals - QQplots show some deviance, shapiro-wilk are significant suggesting lack of normality 
Independence - looks fine
multicolinearity - all within limits
homoscedacity - scale-location plot looks ok



#OUTCOME 5 Life expectancy for women

##Scatter plots

###Life Expectancy ~ GSNI
```{r}
ggplot(analysis17, aes(x = twobias, y = LEbirth2019)) +
  geom_point(color = "violetred3")+
  geom_smooth(method = "lm")+
         labs(x = "percentage with 2 GSNI bias (%)",
                y = "Female life expectancy",
                title = "Scatter plot of GSNI (2 bias) and female life expectancy")

analysis17[,cor.test(twobias,LEbirth2019,method = "spearman")]

```

###Life Expectancy ~ physicians per 1000
```{r}
ggplot(analysis17, aes(x = phy17edit, y = LEbirth2019)) +
  geom_point(color = "violetred3")+
  geom_smooth(method = "lm")+
         labs(x = "physicians per 1000",
                y = "Female life expectancy",
                title = "Scatter plot of physicians per 1000 and female life expectancy")

analysis17[,cor.test(phy17edit,LEbirth2019,method = "spearman")]

```
###Life Expectancy ~ mean years of schooling
```{r}
ggplot(analysis17, aes(x = scl17, y = LEbirth2019)) +
  geom_point(color = "violetred3")+
  geom_smooth(method = "lm")+
         labs(x = "mean years of schooling",
                y = "Female life expectancy",
                title = "Scatter plot of mean years of schooling and female life expectancy")

analysis17[,cor.test(scl17,LEbirth2019,method = "spearman")]

```
###Life Expectancy ~ GDP per capita
```{r}
ggplot(analysis17, aes(x = GDP17, y = LEbirth2019)) +
  geom_point(color = "violetred3")+
  geom_smooth(method = "lm")+
         labs(x = "GDP per capita",
                y = "Female life expectancy",
                title = "GDP per capita and female life expectancy")

analysis17[,cor.test(GDP17,LEbirth2019,method = "spearman")]

```
###Life Expectancy ~ MMR
```{r}
ggplot(analysis17, aes(x = MMR17, y = LEbirth2019)) +
  geom_point(color = "violetred3")+
  geom_smooth(method = "lm")+
         labs(x = "MMR",
                y = "Female life expectancy",
                title = "Scatter plot of MMR and female life expectancy")

analysis17[,cor.test(MMR17,LEbirth2019,method = "spearman")]

```
###Life Expectancy ~ year of index
```{r}
outlier17.3<-analysis17[LEbirth2019<67]
ggplot(analysis17, aes(x = GSNI_PERIOD, y = LEbirth2019)) +
  geom_boxplot(color = "violetred3", notch= TRUE)+
  
         labs(x = "GSNI period",
                y = "Life expectancy at birth (2019)",
                title = "GSNI period of collection and life expectancy")+
ggrepel::geom_text_repel(data = outlier17.3, aes(label = location_name                                        ))  

```


##Univariable regression

###Life Expectancy ~ GSNI
```{r}
m17.17 <- lm(LEbirth2019 ~ twobias, data = analysis17)
summary(m17.17)
plot(m17.17)
```

###Life Expectancy ~ physicians per 1000
```{r}
m17.18 <- lm(LEbirth2019 ~ phy17edit, data = analysis17)
summary(m17.18)
plot(m17.18)
```

###Life Expectancy ~ mean years of schooling
```{r}
m17.19 <- lm(LEbirth2019 ~ scl17, data = analysis17)
summary(m17.19)
plot(m17.19)
```

###Life Expectancy ~ GDP per capita
```{r}
m17.20 <- lm(LEbirth2019 ~ GDP17, data = analysis17)
summary(m17.20)
plot(m17.20)
```
###Life Expectancy ~ MMR
```{r}
m17.21 <- lm(LEbirth2019 ~ MMR17, data = analysis17)
summary(m17.21)
plot(m17.21)
```
###Life Expectancy ~ year of index
```{r}
m17.22 <- lm(LEbirth2019 ~ GSNI_PERIOD, data = analysis17)
summary(m17.22)
plot(m17.22)
```

##Multivariable regression
```{r}

m17.23 <- lm(LEbirth2019 ~ twobias + phy17edit + scl17+ MMR17+ GDP17, data = analysis17)
summary(m17.23)
plot(m17.23)

#added in GSNI PERIOD
m17.24 <- lm(LEbirth2019 ~ twobias + phy17edit + scl17+ MMR17+ GDP17+GSNI_PERIOD, data = analysis17)
summary(m17.24)
plot(m17.24)

#take log of GDP and phy based on scatter plots
m17.242 <- lm(LEbirth2019 ~ twobias + log(phy17edit) + scl17+ MMR17+ log(GDP17)+GSNI_PERIOD, data = analysis17)
summary(m17.242)
plot(m17.242)


car::outlierTest(m17.24)
```


assess for colinearity
```{r}
car::vif(m17.23)
car::vif(m17.24)
#all <4
```

assess for normality of residuals 
```{r}
shapiro.test(m17.23$residuals)


shapiro.test(m17.24$residuals)

#no evidence of non-normality in the residuals
```


```{r}
#check for independence
plot(m17.23$residuals, type = "o")
plot(m17.24$residuals, type = "o")
```


Assumptions
Linearity - residual vs fitted appears within limits (but bit wavy?)
Normality of residuals - QQplots look ok and shapiro wilk not significant
Independence - looks fine
multicolinearity - all within limits
homoscedacity - scale-location plot looks ok around 1

outlier 45

Comparing AIC (using corrected as number of covariates is 10% of data https://stats.stackexchange.com/questions/86768/aicc-for-small-sample-sizes)

```{r}
AICcmodavg::AICc(m17.23)

AICcmodavg::AICc(m17.24) #m24 has lower AICc -> preferable and higher r2
```


#OUTCOME 6 Life expectancy for men

##Scatter plots

###Life Expectancy ~ GSNI
```{r}
ggplot(analysis17, aes(x = twobias, y = MLEbirth2019)) +
  geom_point(color = "violetred3")+
  geom_smooth(method = "lm")+
         labs(x = "percentage with 2 GSNI bias (%)",
                y = "male life expectancy",
                title = "Scatter plot of GSNI (2 bias) and male life expectancy")

analysis17[,cor.test(twobias,MLEbirth2019,method = "spearman")]

```

###Life Expectancy ~ physicians per 1000
```{r}
ggplot(analysis17, aes(x = phy17edit, y = MLEbirth2019)) +
  geom_point(color = "violetred3")+
  geom_smooth(method = "lm")+
         labs(x = "physicians per 1000",
                y = "male life expectancy",
                title = "Scatter plot of physicians per 1000 and male life expectancy")

analysis17[,cor.test(phy17edit,MLEbirth2019,method = "spearman")]

```
###Life Expectancy ~ mean years of schooling
```{r}
ggplot(analysis17, aes(x = scl17, y = MLEbirth2019)) +
  geom_point(color = "violetred3")+
  geom_smooth(method = "lm")+
         labs(x = "mean years of schooling",
                y = "male life expectancy",
                title = "Scatter plot of mean years of schooling and male life expectancy")

analysis17[,cor.test(scl17,MLEbirth2019,method = "spearman")]

```
###Life Expectancy ~ GDP per capita
```{r}
ggplot(analysis17, aes(x = GDP17, y = MLEbirth2019)) +
  geom_point(color = "violetred3")+
  geom_smooth(method = "lm")+
         labs(x = "GDP per capita",
                y = "male life expectancy",
                title = "GDP per capita and male life expectancy")

analysis17[,cor.test(GDP17,MLEbirth2019,method = "spearman")]

```

###Life Expectancy ~ year of index
```{r}

ggplot(analysis17, aes(x = GSNI_PERIOD, y = MLEbirth2019)) +
  geom_boxplot(color = "violetred3", notch= TRUE)+
  
         labs(x = "GSNI period",
                y = "Life expectancy at birth (2019)",
                title = "GSNI period of collection and life expectancy"                                    )

```


##Univariable regression

###Life Expectancy ~ GSNI
```{r}
m17.17m <- lm(MLEbirth2019 ~ twobias, data = analysis17)
summary(m17.17m)
plot(m17.17m)
```

###Life Expectancy ~ physicians per 1000
```{r}
m17.18m <- lm(MLEbirth2019 ~ phy17edit, data = analysis17)
summary(m17.18m)
plot(m17.18m)
```

###Life Expectancy ~ mean years of schooling
```{r}
m17.19m<- lm(MLEbirth2019 ~ scl17, data = analysis17)
summary(m17.19m)
plot(m17.19m)
```

###Life Expectancy ~ GDP per capita
```{r}
m17.20m <- lm(MLEbirth2019 ~ GDP17, data = analysis17)
summary(m17.20m)
plot(m17.20m)
```
###Life Expectancy ~ MMR
```{r}
m17.21m <- lm(MLEbirth2019 ~ MMR17, data = analysis17)
summary(m17.21m)
plot(m17.21m)
```
###Life Expectancy ~ year of index
```{r}
m17.22m <- lm(MLEbirth2019 ~ GSNI_PERIOD, data = analysis17)
summary(m17.22m)
plot(m17.22m)
```

##Multivariable regression
```{r}

m17.23m <- lm(MLEbirth2019 ~ twobias + phy17edit + scl17+  GDP17, data = analysis17)
summary(m17.23m)
plot(m17.23m)

#added in GSNI PERIOD
m17.24m <- lm(MLEbirth2019 ~ twobias + phy17edit + scl17+  GDP17+GSNI_PERIOD, data = analysis17)
summary(m17.24m)
plot(m17.24m)

#take log of GDP and phy based on scatter plots
m17.242m <- lm(MLEbirth2019 ~ twobias + log(phy17edit) + scl17 + log(GDP17) + GSNI_PERIOD, data = analysis17)
summary(m17.242m)
plot(m17.242m)


```

```{r}
#Qatar is again an outlier, removing it doesn't chagne the result
m17.242mnoQ <- lm(MLEbirth2019 ~ twobias + log(phy17edit) + scl17 + log(GDP17) + GSNI_PERIOD, data = noQatar17)
summary(m17.242mnoQ)
plot(m17.242mnoQ)

```



assess for colinearity
```{r}

car::vif(m17.242m)
car::vif(m17.242mnoQ)

#all <4
```



Assumptions
Linearity - residual vs fitted appears within limits (but bit wavy?)
Normality of residuals - QQplots look ok and shapiro wilk not significant
Independence - looks fine
multicolinearity - all within limits
homoscedacity - scale-location plot looks ok around 1

outlier 45

Comparing AIC (using corrected as number of covariates is 10% of data https://stats.stackexchange.com/questions/86768/aicc-for-small-sample-sizes)




#---